### 2025-04-05 geochron@home/manuscript/paper.tex
Added the manuscript

\documentclass[gchron, manuscript]{copernicus}

\usepackage{hyperref,caption}

\begin{document}

\title{FAIR fission track analysis with geochron@home}

\Author[1][p.vermeesch@ucl.ac.uk]{Pieter}{Vermeesch}
\Author[1]{Tim}{Band}
\Author[2]{Jiangping}{He}
\Author[3]{Andrew}{Carter}
\Author[1]{Rex}{Galbraith}

\affil[1]{University College London, London WC1E 6BT, United Kingdom}
\affil[2]{King's College London, London SE1 8WA, United Kingdom}
\affil[3]{Birkbeck, University of London, London WC1E 7HX, United Kingdom}

\runningtitle{FAIR fission track dating with geochron@home}

\runningauthor{P Vermeesch \emph{et al.}}

\received{}
\pubdiscuss{} %% only important for two-stage journals
\revised{}
\accepted{}
\published{}

\firstpage{1}

\maketitle

\begin{abstract}
Fission track geochronology is based on the visual analysis of optical
images. This visual process is prone to observer bias. Fission track
datasets are currently reported as small data tables. The
interpretation of these tables requires a high degree of trust between
the fission track analyst and the user of the data. geochron@home is
software that removes this requirement of trust. It combines a
browser-based `virtual microscope' with an online database to provide
FAIR (Findable, Accessible, Interoperable and Reproducible) access to
fission track data.\medskip

geochron@home serves four different purposes. It can be used (1) to
count fission tracks in `private mode', i.e. hidden from other users
on the internet; (2) to archive fission track images and counts for
inspection by other users; (3) to create tutorials for new students of
the fission track method; and (4) to generate crowd-sourced fission
track data, by serving randomly selected selections of images to
citizen scientists. We illustrate these four applications with
examples that demonstrate (1) geochron@home's ability to compare and
combine fission track counts for multiple users within a lab group;
(2) the use of the geochron@home archive in the peer review system;
(3) the effectiveness of even simple tutorials in teaching novice
users how to count fission tracks; and (4) the `wisdom of crowds' in
fission track identification.\medskip

geochron@home was written in Python and PostgreSQL. Its code is freely
available for inspection and modification, allowing users to set up
their own geochron@home server. Alternatively, users who would like to
upload data to the GaHa, but do not have the facilities to set up
their own server, may use the server at University College London free
of charge. The GaHa accepts image stacks acquired on any type of
digital microscope, and accomodates fission track data (counts and
length measurements) from external fission track analysis suites such
as Fission Track Studio and Track\emph{Flow}.\medskip

We anticipate that the introduction of FAIR workflows will make
fission track data more accurate and more future proof. Storing
fission track data online will benefit future developments in fission
track thermochronology. For example, archival datasets of peer
reviewed fission track counts can be used to train and improve machine
learning algorithms for automated fission track analysis. We hope that
other geochronological methods will follow the fission track
community's lead in FAIR data processing, with benefits to all the
other Earth Science disciplines that depend on geochronological data.
\end{abstract}

\introduction\label{sec:intro}

Science is on an irreversible trajectory towards greater
openness. Geochronology is no exception to this trend, as this journal
illustrates with its open access and review policies. Funding agencies
increasingly demand that research results and data are shared with the
public. Currently, geochronological data are generally provided as
flat tables of dates or isotopic ratio estimates. However, in other
fields of science such as physics, it is common practice to share the
raw unprocessed measurements along with processing instructions
\citep[e.g.,][]{abbott2016}. This paper moves geochronology in the
same direction. It presents a mechanism to generate and store fully
FAIR \citep[Findable, Accessible, Interoperable and
  Reproducible;][]{wilkinson2016} data in the context of fission track
analysis.\medskip

Unlike most other geochronometers, which use mass spectrometers to
estimate parent-daughter ratios, fission tracks are observed under an
optical microscope and counted by a human observer.  In recent years,
digital microscopy has moved fission track data acquisition from the
objective lens of a microscope to the computer screen
\citep{gleadow2009, vanranst2019, gleadow2019}. Ongoing developments
in artificial intelligence generate further opportunities to improve
the throughput and accuracy of fission track data
\citep{nachtergaele2020}. But despite the richness of the digital
datasets produced by these novel tools, fission track data are still
reported as small summary tables. These tables require an unnecessary
degree of trust between the `producer' and `consumer' of the
data.\medskip

geochron@home is a free and open software platform that allows
geochronologists to share their raw fission track data over the
internet for perusal by peer reviewers, colleagues and the general
public. geochron@home is a virtual petrographic microscope connected to a
database with digital image stacks of etched fission track
samples. The platform can be used to acquire, archive and inspect
fission track data in full adherence to the FAIR data principles. In
Section~\ref{sec:architecture}, we will describe geochron@home's software
architecture in five steps. We will show that this architecture
accommodates imagery from any type of digital microscope. It enables
flexible workflows that can be adapted to four different
applications.\medskip

Using image stacks of Mount Dromedary apatite, we will show how
geochron@home can be used to count fission tracks in `private mode'
(Section~\ref{sec:private}); to archive published fission track
datasets in `public mode' (Section~\ref{sec:GaHa}); to build tutorials
for training purposes (Section~\ref{sec:tutorial}); and to
crowd-source fission track data on the internet
(Section~\ref{sec:crowdsourcing}). Because geochron@home is free and
open, it can be extended and improved by any interested party. We make
some suggestions for future improvements in
Section~\ref{sec:outlook}. We hope that the geochron@home's example
will be followed in other geochronological disciplines, as this will
benefit not only geochronology itself, but all the other disciplines
that depend on it (Section~\ref{sec:conclusions}).

\section{Workflow}\label{sec:architecture}

The geochron@home workflow separates the acquisition of
microscope images from their analysis, providing the flexibility to
accommodate data from different microscope manufacturers, which
typically come with their own instrument control software. The
workflow can be broken down into five steps.

\begin{enumerate}
\item Acquisition of z-stacks of microscope images in reflected and
  transmitted light for each of the grains in a sample and,
  optionally, for the accompanying external detector
  (Figure~\ref{fig:EDM}). At University College London, this first
  step is currently accomplished by a Python macro within Zeiss' Zen
  Blue software.  However, geochron@home can also accommodate imagery
  from other platforms, such as Fission Track Studio
  \citep[Zeiss;][]{gleadow2009} and Track\emph{Flow}
  \citep[Nikon;][]{vanranst2019}.

\item Prepare the z-stacks for uploading to the geochron@home
  database. This database requires that the images are organised as a
  nested sequence of directories, in which a `project' consists of
  `samples' that comprise a number of `grains'.  Each grain
  corresponds to a numbered sequence of \texttt{.jpeg} images, one for
  each layer of the z-stack. Note that the raw microscope images are
  generally not stored as \texttt{.jpeg} files but in uncompressed
  \texttt{.czi} (for Zen Blue), \texttt{.tif} (for Fission Track
  Studio) or \texttt{.nd2} (for Nikon/Track\emph{Flow})
  formats. Conversion from these raw images to sequences of
  \texttt{.jpeg} files is done with a shell script using Imagemagick
  \citep{still2006}.\medskip

In addition to the sequence of \texttt{.jpeg} images, a low level
`grain' folder can also include an optional file called
\texttt{roi.json} containing the vertices of a default region of
interest for spontaneous (and/or induced) track counting. The database
structure can also accommodate existing fission track counting
results.  For example, if a user has already counted their fission
tracks in Fission Track Studio, then they can store those results in a
\texttt{.json} file at the `sample' level directory. Because Fission
Track Studio stores its results in an \texttt{.xml} format, a second
conversion script is needed to translate those results into an
equivalent \texttt{.json} format.

\item Upload the data to the geochron@home platform. geochron@home is
  a Django web-app with a PostgreSQL database. The database can be
  accessed via a Python API and a (more limited) web-based GUI
  interface. Accessing the API requires administrator
  privileges. Administrators can create projects, samples and grains;
  download data; and set the access rights of `ordinary' users.
  Projects can be private or public. The source code and installation
  instructions for geochron@home are freely available over GitHub (see
  the Data Availability statement at the end of this paper). This
  allows fission track users to set up their own
  server. Alternatively, fission track labs can upload their data to
  the UCL server by contacting the corresponding author.

\item Analyse the images with a browser-based `virtual fission track
  microscope'. This virtual instrument has a simple user interface
  with controls to zoom, pan and focus in or out of the digital image
  stack. Depending on the permissions granted to the user by the
  administrator, the virtual microsocope offers a number of different
  options.  Entry level `ordinary' users are only allowed to count
  tracks by clicking within a pre-defined `region of interest'. In
  contrast, `superusers' are allowed to define their own regions of
  interest. Once the user is satisfied that they have counted all the
  fission tracks in a particular grain, they can submit the results to
  the server. They are then presented with a new set of images until
  all grains are counted.

\item Post-processing. The fission track data can either be downloaded
  as a flat data table of counts and area, or as a \texttt{.json} file
  containing the locations of all the coounted
  tracks. geochron@home does not provide any tools to
  post-process these files. They are meant to be passed on to other
  tools such as spreadsheet applications or IsoplotR
  \citep{vermeesch2018c}.

\end{enumerate}

The five-step workflow can be used for several applications, including
(1) conventional fission track analysis; (2) archiving published
fission track results; (3) building tutorials; and (4) crowd-sourcing
fission track data. The next sections of this paper will illustrate
these four different applications with real world examples.\medskip

%\begin{figure}[!ht]
{ \centering \includegraphics[width=12cm]{EDM.jpg}
  \captionof{figure}{Screenshots of raw fission track data in
    geochron@home. GR: an apatite grain in reflected light; GT: the
    same grain in transmitted light; MR: the corresponding mica
    detector in reflected light; MT: the mica detector in transmitted
    light. White rectangles mark the region of interest (ROI), within
    which an analyst has counted fission tracks by marking their etch
    pits (shown in yellow). The raw data for this figure can be viewed
    on the geochron@home archive
    (\url{https://github.com/pvermees/GaHa}).\medskip}
  \label{fig:EDM}
}%\end{figure}

\section{Counting fission tracks in `private mode'}\label{sec:private}

Administrators can define regions of interest (ROI) and count or edit
the fission track coordinates of any grain in a geochron@home
database. They can also assign other users to groups, and give these
groups access to a subset of the projects in the
database. Administrators have fine control over the permissions of the
groups. For example, they can allow the members of a group to count
fission tracks in the predefined ROIs, or to define their own
ROIs. Groups provide a mechanism to compare and combine the results of
multiple analysts of the same sample. Figure~\ref{fig:AvP} illustrates
this with two sets of fission track density estimates (in tracks per
\textmu{m}\textsuperscript{2}) for the same sample of Mount Dromedary
apatite \citep{green1985b}, analysed by two users (PV and AC).\medskip

The two analysts defined their own ROIs, resulting in slightly
different density estimates
(Figure~\ref{fig:AvP}a). Figure~\ref{fig:AvP}b compares the track
density estimates of the two analysts for 25 apatite grains on a
radial plot, using the assumption that the density estimates are
statistically independent from each other. Let $N_{PV}$ and $N_{AC}$
be the number of spontaneous tracks counted by PV and AC over areas
$A_{PV}$ and $A_{AC}$, respectively. Then Figure~\ref{fig:AvP}b
assumes that the ratio of the estimated track densities
$\hat{\rho}_{PV} = N_{PV}/A_{PV}$ and $\hat{\rho}_{AC} =
N_{AC}/A_{AC}$ is associated with the following uncertainty:
\begin{equation}
  \hat{\sigma} =
  \sqrt{\mbox{var}\left(\frac{\hat{\rho}_{PV}}{\hat{\rho}_{AC}}\right)} =
  \frac{\hat{\rho}_{PV}}{\hat{\rho}_{AC}} \, \sqrt{\frac{1}{N_{PV}} +
    \frac{1}{N_{PV}} }
  \label{eq:pessimistic}
\end{equation}

In reality, the uncertainties are strongly correlated due to the
significant overlap between the counting areas chosen by the two
analysts. This overlap causes the density ratios to be underdispersed
with respect to the assigned uncertainties ($\mbox{MSWD}=0.34$ in
Figure~\ref{fig:AvP}b).  Equation~\ref{eq:pessimistic} can be modified
to remove the error correlation under the assumption that the two
analysts counted exactly the same features in the overlapping area
($A_0$):
\begin{equation}
  \hat{\sigma}' =
  \frac{\hat{\rho}_{PV}}{\hat{\rho}_{AC}}
  \sqrt{\frac{A_{PV}+A_{AC}}{N_{PV}+N_{AC}}}\,
  \sqrt{
    \frac{1}{A_{PV}} + \frac{1}{A_{AC}} - \frac{2\,A_0}{A_{PV}\,A_{AC}}
  }
  \label{eq:optimistic}
\end{equation}

If there is no overlap between the two counting areas (i.e., if
$A_0=0$), then $\hat{\sigma}' \approx \hat{\sigma}$. Conversely, if
the counting areas are identical (i.e., if $A_{PV}=A_{AC}=A_0$), and
if the two analysts produce perfectly consistent counts, then
$\hat{\sigma}'=0$.\medskip

Equation~\ref{eq:optimistic} produces a radial plot
(Figure~\ref{fig:AvP}c) that is overdispersed ($\mbox{MSWD}=4.2$) with
respect to $\hat{\sigma}'$. This suggests that the assumption of
perfect consistency is also wrong. Detailed inspection of the counted
features in the virtual microscope reveals that the two analysts are
in perfect agreement for long fission tracks, but that they
occasionally disagree about the shorter features. The ratio of the
pooled track densities indicates that PV undercounts the Mount
Dromedary apatite by 7\% relative to AC.
\medskip

In summary, it could be said that Equation~\ref{eq:pessimistic}
represents an overly `pessimistic' view of analysts' abilities to
reproduce each other's results, whereas Equation~\ref{eq:optimistic}
is too `optimistic'. The truth lies between the two scenarios.  The
existence of `observer bias' is well documented \citep{tamer2025}.  It
is one of the reasons why fission track analysis is often done
relative to age standards: observer bias does not have to be a problem
provided that it is consistent between grains, and between
samples.\medskip

%\begin{figure}[!ht]
{ \centering \includegraphics[width=\linewidth]{AvP.pdf}
  \captionof{figure}{Comparison of fission track data for Mount
    Dromedary apatite by two analysts (PV = Pieter Vermeesch and AC =
    Andrew Carter). a) the track counts and ROIs of both users for
    grain 4648, shown in blue (PV) and red (AC); b) radial plot for
    the ratio of the track densities of both users for all 25 grains
    analysed by the two analysts, using the `pessimistic' scenario of
    Equation~\ref{eq:pessimistic}; c) radial plot of the same data,
    using the `optimistic' error propagation formula of
    Equation~\ref{eq:optimistic}.
    \medskip}
  \label{fig:AvP}
}%\end{figure}

\section{The geochron@home archive (GaHa)}\label{sec:GaHa}

Once a set of fission track images has been analysed and the analyst
is confident that the results are accurate, the status of the results
can be changed from private to public. This makes the results visible
over the internet as a list of URLs, where each grain number and user
ID corresponds to a unique address. The geochron@home archive
(GaHa) brings fission track geochronology into the era of FAIR
science. It allows peer reviewers to inspect the raw data from which
thermochronological inferences are made. The archive is open to
submissions from any fission track laboratory, free of charge.
However, as mentioned in Section~\ref{sec:architecture}, it is also
possible to establish a new archive elsewhere. At the time of writing,
GaHa contains data for three studies:

\begin{enumerate}
\item{\citet{guo2025}}: this is an LA-ICP-MS based fission track study
  of detrital apatite from the northeastern Tibetan Plateau. It
  contains image stacks of semitracks in 1146 apatite grains from 16
  different samples.
\item{\citet{tamer2025}}: this is a round-robin study in which digital
  image stacks of 44 apatite crystals were circulated among 14
  different analysts. These analysts used the FastTrack image analysis
  software (which is part of the Fission Track Studio suite) to define
  their own ROIs and count the semitracks and horizontally confined
  fission tracks in them. GaHa presents the results of the round-robin
  experiment as a ${44}\times{14}$ grid of URLs.
\item{This study}: All the raw fission track data used in this article
  are available on GaHa, along with the post-processing software that
  was used to produce the figures. Together, these resources provide
  the reader with all the information needed to fully reproduce our
  results, `from cradle to grave'. To our knowledge, this is the first
  geochronological study to do so.
\end{enumerate}

\section{geochron@home tutorials}\label{sec:tutorial}

Given the right permissions (assigned by an administrator), users can
build tutorial pages by annotating features in fission track images.
These features can be tracks or other objects such as scratches,
inclusions, dislocations or holes. A selection of tutorial pages is
presented to new users when they first log into
geochron@home. They must complete the tutorial before being
allowed to count fission tracks.  The tutorial pages can be revisited
at any time by visiting the corresponding link on the
geochron@home landing page.\medskip

The annotations in the current tutorial pages were made by an
experienced fission track analyst (Andrew Carter). They provide a
quick and easy mechanism to train novice users in the art of fission
track analysis. At the time of writing, the tutorial pages offer only
a limited degree of interactivity. Users can click on features to read
the annotations. However, future plans include the addition of fully
interactive `quizzes' (see Section~\ref{sec:conclusions}). Despite the
simplicity of the current tutorials, they are already quite effective,
as demonstrated by the crowd-sourcing experiment described in the next
section of this paper.

\section{Crowd-sourcing fission track data}\label{sec:crowdsourcing}

%\begin{figure}[!ht]
{ \centering \includegraphics[width=15cm]{radialcrowd.pdf}
  \captionof{figure}{a) box plots of 25 crowd-sourced sets of fission
    track counts for 70 analysts. b) radial plot comparing the paired
    track counts for the first two grains in the dataset (4649 and
    4673). The estimate of user PV is shown in blue. Analyst AC used a
    different ROI and is not shown.\medskip}
  \label{fig:radialcrowd}
}%\end{figure}

In 1906, Sir Francis Galton visited a county fair in which a contest
was held to guess the weight of an ox. 787 people participated in the
event. Galton discovered that the median of all their estimates was
within 0.8\% of the true weight of the ox and more accurate than 90\%
of the individual estimates. Such is the `wisdom of crowds'
\citep{galton1907a}. Similar effects are seen in fission track
geochronology.\medskip

A 1984 interlaboratory comparison study \citep{miller1985} showed that
the average of several fission track age estimates is closer to the
known age of mineral standards than the age obtained by any individual
observer. Unfortunately, routine measurement of fission track samples
by multiple analysts is prohibitively expensive in a normal laboratory
environment. geochron@home changes this, by bringing fission
track dating to a proverbial `county fair' of citizen
scientists.\medskip

Figure~\ref{fig:crowdsourcing} presents the results of a
crowd-sourcing experiment that was carried out at UCL as part of an
undergraduate course in isotope geology. 70~students were asked to
create an account on the geochron@home server by selecting a
unique username and password. After spending a few minutes to complete
the compulsory tutorial (Section~\ref{sec:tutorial}), the students
were asked to count fission tracks in randomly assigned image stacks
of Mount Dromedary apatite. Each student was required to analyse at
least 15 grains, using pre-defined ROIs. In a matter of hours, the
students amassed a dataset of 37,331 fission track counts in 34
grains. This large dataset was parsed into separate data files --one
for each student-- which they processed during an assessed practical
exercise.\medskip

The raw counts are shown as a series of box plots in
Figure~\ref{fig:crowdsourcing}a for the same selection of 25 grains
that were analysed by AC and PV in Section~\ref{sec:private}. The box
plots exhibit significant variability between the grains and between
the students. This inter-grain variability is expected. It reflects
the different sizes of the ROIs, and the variable U-content of the
grains. The variability of the track counts for each grain results
from observer bias. This phenomenon was previously discussed in
Section~\ref{sec:private}. However, the bias is much stronger for the
crowd-sourced dataset than it was for the two `expert' analysts of
Section~\ref{sec:private}.\medskip

Figure~\ref{fig:crowdsourcing}b compares the 42 raw fission track
counts of the first two grains on a radial plot.  The pairwise ratios
of the counts for each analyst are more consistent than the raw counts
of Figure~\ref{fig:crowdsourcing}a.  Students who over- or undercount
the first grain, tend to over- or undercount the second grain as
well. Nevertheless, the results are overdispersed with respect to the
counting uncertainties, even when using the `pessimistic' scenario of
Equation~\ref{eq:pessimistic}.  The poor consistency of the results
reflects the very limited training received by the students. But
despite these inconsistencies, the opinion of an `expert' (PV, shown
in blue) falls the middle of the estimates obtained by the citizen
scientists. This suggests that the collective wisdom of the entire
group of students is greater than that of any individual
analyst.\medskip

As metioned in Section~\ref{sec:architecture}, geochron@home stores
the actual track positions marked by the users. This raw data can be
downloaded as a \texttt{.json} file and inspected in detail.
Figure~\ref{fig:radialcrowd} is a two-dimensional histogram for the
x-y positions of all the track positions generated by the students in
the two apatite crystals. Visual comparison of the histogram with the
optical image confirms that the students unanimously identified the
most obvious semitracks, which contain a clearly visible etch pit and
tail. Shorter and fainter tracks received fewer clicks. Datasets like
this can be used to replace integer counts of fission tracks with
probabilities, reflecting the ambiguity of some fission track
datasets.\medskip

Figure~\ref{fig:radialcrowd} also shows that some students counted the
tails of the fission tracks rather than their etch pits, despite being
told the opposite in the tutorial. Fixing this issue will require some
improvements to the tutorial pages (Section~\ref{sec:outlook}).

%\begin{figure}[!ht]
{  \centering
  \includegraphics[width=9cm]{4649vs4673.pdf}
  \captionof{figure}{
    a,b) Optical images in transmitted light of 
    Mount Dromedary apatites 4649 and 4673 in the crowd-sourcing experiment.
    c,d) Two-dimensional histogram of the track locations for the
    two grains, as identified by the citizen scientists.
    \medskip}
  \label{fig:crowdsourcing}
}%\end{figure}

\section{Outlook}\label{sec:outlook}

geochron@home has been in development for 10 years and
remains a work in progress. Planned improvements include:

\begin{enumerate}
\item Interactive tutorial pages. To ensure that novice users do not
  count the tails but the etch pits of fission tracks, we will add a
  `quiz' to the tutorial pages. Only users who click on the `correct'
  features in an unlabelled set of images will be allowed to count new
  samples.
\item Length measurements. The geochron@home archive already
  contains horizontally confined fission track measurements
  \citep{tamer2025}.  However, these results must be generated
  externally (e.g., using Fission Track Studio) and uploaded via a
  \texttt{.json} file.  The virtual microscope curently lacks the
  functionality to generate length data within
  geochron@home. This functionality will be added in a future
  update.
\item Dpar and Dper. Etch pits are currently stored as simple sets of
  x- and y-coordinates.  In reality, etch pits have a finite length
  (`Dpar') and width (`Dper'), which serve as useful indicators for
  the horizontal etch rates along the c-axis and parallel to it
  \citep{donelick1993}. Functions will be added measure and visualise
  these type of data in geochron@home.
\item Machine learning. Data science is experiencing an artificial
  intelligence revolution that has already started to transform the
  fission track method \citep{nachtergaele2020}.  Convolutional neural
  networks must be trained with example data. geochron@home
  is ideally suited for this task. Section~\ref{sec:crowdsourcing}
  showed how the collective wisdom of multiple fission track analysts
  can label fission track images with probabilities rather than
  counts. This data format is close to the form in which data are
  treated within an AI algorithm.\medskip

  Once trained on historical data, AI algorithms can be used to count
  fission tracks automatically. Following the model of Fission Track
  Studio, machine learning can be used to reverse the fission track
  counting process. Instead of asking users to count the fission
  tracks in a sample, the software can ask them to check the results
  proposed by an AI algorithm, and to remove any features that are
  \emph{not} fission tracks.\medskip

  Regardless of whether fission tracks were counted manually or with a
  machine, the value of the geochron@home archive remains the
  same. It is important to document data so that samples can be
  reanalysed in the future, for example when a new and improved
  generation of machine learning algorithms becomes available.
\end{enumerate}

These four improvements will be made by ourselves pending additional
funding. However, because geochron@home is free and open
software, we invite any interested parties to join the effort and
extend or improve our code.

\section{Conclusions}\label{sec:conclusions}

This paper introduced geochron@home, a software platform for
FAIR fission track analysis. We demonstrated four different
applications of this platform using real data. Putting the FAIR data
paradigm into practice, all the imagery, counts and source code for
this paper are publicly available. Using these resources, the reader
can reproduce all the results that were presented in this publication.
We hope that other geochronologists will follow this example.  FAIR
data promises to address the reproducibility crisis in science
\citep{miyakawa2020}.\medskip

geochron@home's rich archive of raw data can be reanalysed in
the future. We anticipate that the adoption of FAIR data processing
workflows will open up new research opportunities. For example,
archived pairs of peer-reviewed fission track images and counts could
be used to train the next generation of automated machine learning
algorithms. Conversely, it is also possible that future improvements
in fission track images analysis will be used to update the count data
for published datasets, improved their accuracy.\medskip

Another advantage of the geochron@home wokflow is the
separation of image acquisition and image analysis. This separation
reduces the hardware requirements for fission track dating. It opens
up the possibility to share resources. State-of-the-art digital
microscopes are expensive. Using geochron@home, a single
microscope can serve multiple users and make fission track analysis
more affordable.\medskip

The fission track method has always been a test bed for new
geochronological developments.  Because fission track data are
imprecise, the fission track community has sollicited the help of
statisticians and mathematicians to develop its analytical protocols.
Other geochronological communities are still catching up with concepts
and tools such as overdispersion, mixture modelling and radial plots,
which have been commonplace in fission track dating for decades
\citep{vermeesch2019b}. In a similar vein, the subjective nature of
fission track identification has prompted the fission track community
to organise inter-laboratory comparisons and round-robin studies long
before other geochronological communities
\citep{miller1985,tamer2025}.\medskip

With the development of geochron@home, fission tracks is once
again ahead of the pack in terms of FAIR data analysis.
geochron@home currently only stores images and counts. This
is enough to reproduce the results of fission track studies using the
external detector method, but not for LA-ICP-MS based data. FAIR data
processing of LA-ICP-MS data requires a new generation of mass
spectrometer data reduction software. We are currently working on
this. The development of FAIR ICP-MS data pipelines will not only
benefit fission track dating but other chronometers as well, such as
in-situ U--Pb, Rb--Sr and Lu--Hf \citep{vermeesch2025}.\medskip

With the establishment of FAIR data, the era of `fiction tracks' will
be gone for good, and geochronology will be well placed to avoid the
reproducibility problems that have plagued other fields of science.

\codedataavailability{geochron@home is free software released under
  the GPL-3 license. The package and its source code are available
  from \url{https://github.com/pvermees/geochron-at-home} (last
  access: April~2, 2025).  The raw data (imagery) are available at the
  geochron@home archive (\url{https://github.com/pvermees/GaHa}, last
  access: April~3, 2025).  R-scripts to reproduce the figures are
  provided in the supplementary information
  (\url{https://github.com/pvermees/supplements}, last access:
  April~3, 2025).}

%\appendix
%\section{}

%\appendix{}

%\noappendix

\authorcontribution{Pieter Vermeesch designed the study, acquired the
  funding, counted fission tracks and wrote the paper. Jiangping He
  created geochron@home. Tim Band expanded geochron@home and wrote the
  accompanying microscope image acquisition software. Andrew Carter
  provided the samples, prepared the training data and counted fission
  tracks. Rex Galbraith derived Equation~\ref{eq:optimistic}.}

\competinginterests{PV is an Associate Editor of \emph{Geochronology}.}

\begin{acknowledgements}
This research has been supported by the Natural Environment Research
Council (grant no. NE/T001518/1), awarded to Pieter Vermeesch. We would like
to thank the students of GEOL0017 (`Isotope Geology') for their contribution
to the crowd-sourcing experiment of Section~\ref{sec:crowdsourcing}.
\end{acknowledgements}

\bibliographystyle{copernicus}
\bibliography{biblio.bib}

\end{document}

### 2025-05-09 geochron@home/manuscript/paper.tex
Updated g@h code and manuscript

@@
 
-\runningtitle{FAIR fission track dating with geochron@home}
+\runningtitle{FAIR fission track analysis with geochron@home}
 
@@
 
-geochron@home was written in Python and PostgreSQL. Its code is freely
+geochron@home was written in Python and Javascript. Its code is freely
 available for inspection and modification, allowing users to set up
 their own geochron@home server. Alternatively, users who would like to
-upload data to the GaHa, but do not have the facilities to set up
+upload data to the archive, but do not have the facilities to set up
 their own server, may use the server at University College London free
-of charge. The GaHa accepts image stacks acquired on any type of
+of charge. The archive accepts image stacks acquired on any type of
 digital microscope, and accomodates fission track data (counts and
@@
 other geochronological methods will follow the fission track
-community's lead in FAIR data processing, with benefits to all the
+community's lead in FAIR data processing. This would benefit all the
 other Earth Science disciplines that depend on geochronological data.
@@
 
-Unlike most other geochronometers, which use mass spectrometers to
+Unlike most other geochronometers, which require mass spectrometers to
 estimate parent-daughter ratios, fission tracks are observed under an
@@
 
-The geochron@home workflow separates the acquisition of
-microscope images from their analysis, providing the flexibility to
-accommodate data from different microscope manufacturers, which
-typically come with their own instrument control software. The
-workflow can be broken down into five steps.
+The geochron@home workflow separates the acquisition of microscope
+images from their analysis, providing the flexibility to accommodate
+data from different microscope manufacturers. The workflow can be
+broken down into five steps.
 
@@
   \texttt{roi.json} containing the vertices of a default region of
-interest for spontaneous (and/or induced) track counting. The database
-structure can also accommodate existing fission track counting
-results.  For example, if a user has already counted their fission
-tracks in Fission Track Studio, then they can store those results in a
-\texttt{.json} file at the `sample' level directory. Because Fission
-Track Studio stores its results in an \texttt{.xml} format, a second
-conversion script is needed to translate those results into an
-equivalent \texttt{.json} format.
+  interest for spontaneous (and/or induced) track counting. The
+  database structure can also accommodate existing fission track
+  counting results.  For example, if a user has already counted their
+  fission tracks in Fission Track Studio, then they can store those
+  results in a \texttt{.json} file at the `sample' level
+  directory. Because Fission Track Studio stores its results in an
+  \texttt{.xml} format, a second conversion script is needed to
+  translate those results into an equivalent \texttt{.json} format.
 
 \item Upload the data to the geochron@home platform. geochron@home is
-  a Django web-app with a PostgreSQL database. The database can be
-  accessed via a Python API and a (more limited) web-based GUI
-  interface. Accessing the API requires administrator
-  privileges. Administrators can create projects, samples and grains;
-  download data; and set the access rights of `ordinary' users.
-  Projects can be private or public. The source code and installation
-  instructions for geochron@home are freely available over GitHub (see
-  the Data Availability statement at the end of this paper). This
-  allows fission track users to set up their own
-  server. Alternatively, fission track labs can upload their data to
-  the UCL server by contacting the corresponding author.
+  a Django web-app featuring a user interface powered by the Leaflet
+  library (JavaScript).  The database can be accessed via a Python API
+  and a (more limited) web-based GUI. Accessing the API requires
+  administrator privileges. Administrators can create projects,
+  samples and grains; download data; and set the access rights of
+  `ordinary' users.  Projects can be private or public. The source
+  code and installation instructions for geochron@home are freely
+  available over GitHub (see the Data Availability statement at the
+  end of this paper). This allows fission track users to set up their
+  own server. Alternatively, fission track labs can upload their data
+  to the UCL server by contacting the corresponding author.
 
@@
   as a flat data table of counts and area, or as a \texttt{.json} file
-  containing the locations of all the coounted
-  tracks. geochron@home does not provide any tools to
-  post-process these files. They are meant to be passed on to other
-  tools such as spreadsheet applications or IsoplotR
-  \citep{vermeesch2018c}.
+  containing the locations of all the counted tracks. geochron@home
+  does not provide any tools to post-process these files. They are
+  meant to be passed on to other tools such as spreadsheet
+  applications or IsoplotR \citep{vermeesch2018c}.
 
@@
 { \centering \includegraphics[width=12cm]{EDM.jpg}
-  \captionof{figure}{Screenshots of raw fission track data in
-    geochron@home. GR: an apatite grain in reflected light; GT: the
-    same grain in transmitted light; MR: the corresponding mica
-    detector in reflected light; MT: the mica detector in transmitted
-    light. White rectangles mark the region of interest (ROI), within
-    which an analyst has counted fission tracks by marking their etch
-    pits (shown in yellow). The raw data for this figure can be viewed
-    on the geochron@home archive
+  \captionof{figure}{
+    Screenshots of raw fission track data for the external detector
+    method in geochron@home. GR: an apatite grain in reflected light;
+    GT: the same grain in transmitted light; MR: the corresponding
+    mica detector in reflected light; MT: the mica detector in
+    transmitted light. White rectangles mark the region of interest
+    (ROI), within which an analyst has counted fission tracks by
+    marking their etch pits (shown in yellow). The raw data for this
+    figure can be viewed on the geochron@home archive
     (\url{https://github.com/pvermees/GaHa}).\medskip}
@@
 database. Administrators have fine control over the permissions of the
-groups. For example, they can allow the members of a group to count
-fission tracks in the predefined ROIs, or to define their own
-ROIs. Groups provide a mechanism to compare and combine the results of
-multiple analysts of the same sample. Figure~\ref{fig:AvP} illustrates
-this with two sets of fission track density estimates (in tracks per
-\textmu{m}\textsuperscript{2}) for the same sample of Mount Dromedary
-apatite \citep{green1985b}, analysed by two users (PV and AC).\medskip
+groups. For example, they can allow the members of one group to define
+their own ROIs, whilst requiring members of another group to count
+fission tracks in predefined ROIs.  Groups provide a mechanism to
+compare and combine the results of multiple analysts of the same
+sample. Figure~\ref{fig:AvP} illustrates this with two sets of fission
+track density estimates (in tracks per \textmu{m}\textsuperscript{2})
+for the same sample of Mount Dromedary apatite \citep{green1985b},
+analysed by two users (PV and AC).\medskip
 
 The two analysts defined their own ROIs, resulting in slightly
-different density estimates
-(Figure~\ref{fig:AvP}a). Figure~\ref{fig:AvP}b compares the track
-density estimates of the two analysts for 25 apatite grains on a
-radial plot, using the assumption that the density estimates are
-statistically independent from each other. Let $N_{PV}$ and $N_{AC}$
-be the number of spontaneous tracks counted by PV and AC over areas
-$A_{PV}$ and $A_{AC}$, respectively. Then Figure~\ref{fig:AvP}b
-assumes that the ratio of the estimated track densities
-$\hat{\rho}_{PV} = N_{PV}/A_{PV}$ and $\hat{\rho}_{AC} =
-N_{AC}/A_{AC}$ is associated with the following uncertainty:
+different density estimates (Figure~\ref{fig:AvP}a) for 25 apatite
+grains.  Let $N_{PV}$ and $N_{AC}$ be the number of spontaneous tracks
+counted by PV and AC over areas $A_{PV}$ and $A_{AC}$,
+respectively. Then the ratio of the estimated track densities are
+given by $\hat{\rho}_{PV} = N_{PV}/A_{PV}$ and $\hat{\rho}_{AC} =
+N_{AC}/A_{AC}$. The uncertainties of these density ratios are strongly
+correlated due to the significant overlap between the counting areas
+chosen by the two analysts:
 \begin{equation}
-  \hat{\sigma} =
-  \sqrt{\mbox{var}\left(\frac{\hat{\rho}_{PV}}{\hat{\rho}_{AC}}\right)} =
-  \frac{\hat{\rho}_{PV}}{\hat{\rho}_{AC}} \, \sqrt{\frac{1}{N_{PV}} +
-    \frac{1}{N_{PV}} }
-  \label{eq:pessimistic}
+  \frac{\mbox{se}(\hat{\rho}_{PV}/\hat{\rho}_{AC})}{\hat{\rho}_{PV}/\hat{\rho}_{AC}}
+  \approx
+  \sqrt{
+    \frac{1}{N_{PV}} + \frac{1}{N_{AC}} - \frac{2\,N_{0}}{N_{PV}\,N_{AC}}
+  }
+  \label{eq:NO}
 \end{equation}
 
-In reality, the uncertainties are strongly correlated due to the
-significant overlap between the counting areas chosen by the two
-analysts. This overlap causes the density ratios to be underdispersed
-with respect to the assigned uncertainties ($\mbox{MSWD}=0.34$ in
-Figure~\ref{fig:AvP}b).  Equation~\ref{eq:pessimistic} can be modified
-to remove the error correlation under the assumption that the two
-analysts counted exactly the same features in the overlapping area
-($A_0$):
-\begin{equation}
-  \hat{\sigma}' =
-  \frac{\hat{\rho}_{PV}}{\hat{\rho}_{AC}}
-  \sqrt{\frac{A_{PV}+A_{AC}}{N_{PV}+N_{AC}}}\,
-  \sqrt{
-    \frac{1}{A_{PV}} + \frac{1}{A_{AC}} - \frac{2\,A_0}{A_{PV}\,A_{AC}}
-  }
-  \label{eq:optimistic}
-\end{equation}
+\noindent where $N_{0}$ is the number of tracks counted by both
+observers in the area of overlap ($A_0$, say) between their respective
+ROIs.  If there is no overlap between the two counting areas (i.e., if
+$A_0=0$), then Equation~\ref{eq:NO} reduces to the conventional
+coefficient of variation for the ratio of two independent Poisson
+variables \citep{galbraith1990a}.  Conversely, if the counting areas
+are identical (i.e., if $A_{PV}=A_{AC}=A_0$), and if the two analysts
+produce perfectly consistent counts (i.e. if $N_{PV} = N_{AC} =
+N_{0}$), then Equation~\ref{eq:NO} returns a zero value.\medskip
 
-If there is no overlap between the two counting areas (i.e., if
-$A_0=0$), then $\hat{\sigma}' \approx \hat{\sigma}$. Conversely, if
-the counting areas are identical (i.e., if $A_{PV}=A_{AC}=A_0$), and
-if the two analysts produce perfectly consistent counts, then
-$\hat{\sigma}'=0$.\medskip
-
-Equation~\ref{eq:optimistic} produces a radial plot
-(Figure~\ref{fig:AvP}c) that is overdispersed ($\mbox{MSWD}=4.2$) with
-respect to $\hat{\sigma}'$. This suggests that the assumption of
-perfect consistency is also wrong. Detailed inspection of the counted
-features in the virtual microscope reveals that the two analysts are
-in perfect agreement for long fission tracks, but that they
-occasionally disagree about the shorter features. The ratio of the
-pooled track densities indicates that PV undercounts the Mount
-Dromedary apatite by 7\% relative to AC.
-\medskip
-
-In summary, it could be said that Equation~\ref{eq:pessimistic}
-represents an overly `pessimistic' view of analysts' abilities to
-reproduce each other's results, whereas Equation~\ref{eq:optimistic}
-is too `optimistic'. The truth lies between the two scenarios.  The
-existence of `observer bias' is well documented \citep{tamer2025}.  It
-is one of the reasons why fission track analysis is often done
-relative to age standards: observer bias does not have to be a problem
-provided that it is consistent between grains, and between
-samples.\medskip
+The radial plot of Figure~\ref{fig:AvP}b contains two sets of counts,
+with $\sum N_{PV} = 686$ and $\sum N_{AC} = 679$, and a weighted mean
+track density ratio $\hat{\rho}_{PV}/\hat{\rho}_{PV} = 0.94$ with
+relative standard error 0.013. PV counted 600 tracks in those common
+areas $A_0$, of which 549 were also counted by AC. Conversely, AC
+counted 622 tracks of which 549 were also counted by PV. The ratio of
+the two analysts' track density estimates based just on the common
+area is therefore 600/622 = 0.965 with relative standard error 0.016,
+calculated from Equation~\ref{eq:NO}. This is close to the weighted
+mean ratio of 0.94 (Figure~\ref{fig:AvP}b), and is slightly nearer to
+1. It indicates that PV under-counts the Mount Dromedary apatite by
+3.5\% relative to AC.  The existence of `observer bias' is well
+documented \citep{tamer2025}.  It is one of the reasons why fission
+track analysis is often done relative to age standards: observer bias
+does not have to be a problem provided that it is consistent between
+grains, and between samples.\medskip
 
 %\begin{figure}[!ht]
-{ \centering \includegraphics[width=\linewidth]{AvP.pdf}
+{ \centering \includegraphics[width=12cm]{AvP.pdf}
   \captionof{figure}{Comparison of fission track data for Mount
@@
     the ratio of the track densities of both users for all 25 grains
-    analysed by the two analysts, using the `pessimistic' scenario of
-    Equation~\ref{eq:pessimistic}; c) radial plot of the same data,
-    using the `optimistic' error propagation formula of
-    Equation~\ref{eq:optimistic}.
+    analysed by the two analysts, using Equation~\ref{eq:NO}.
     \medskip}
@@
   different analysts. These analysts used the FastTrack image analysis
-  software (which is part of the Fission Track Studio suite) to define
-  their own ROIs and count the semitracks and horizontally confined
-  fission tracks in them. GaHa presents the results of the round-robin
-  experiment as a ${44}\times{14}$ grid of URLs.
+  software \citep[which is part of the Fission Track Studio
+    suite;][]{gleadow2009} to define their own ROIs and count the
+  semitracks and horizontally confined fission tracks in them. GaHa
+  presents the results of the round-robin experiment as a
+  ${44}\times{14}$ grid of URLs.
 \item{This study}: All the raw fission track data used in this article
@@
 The annotations in the current tutorial pages were made by an
-experienced fission track analyst (Andrew Carter). They provide a
-quick and easy mechanism to train novice users in the art of fission
-track analysis. At the time of writing, the tutorial pages offer only
-a limited degree of interactivity. Users can click on features to read
-the annotations. However, future plans include the addition of fully
-interactive `quizzes' (see Section~\ref{sec:conclusions}). Despite the
-simplicity of the current tutorials, they are already quite effective,
-as demonstrated by the crowd-sourcing experiment described in the next
-section of this paper.
+experienced fission track analyst (Andrew Carter). This basic tutorial
+provides a quick and easy mechanism to train novice users in the art
+of fission track analysis. The tutorial pages are in their infancy,
+and offer only a limited degree of interactivity. Users can click on
+features to read the annotations. The tutorial will grow with input
+from experienced analysts and future plans include the addition of
+fully interactive `quizzes' (Section~\ref{sec:conclusions}). Despite
+the simplicity of the current tutorials, they are already quite
+effective, as demonstrated by the crowd-sourcing experiment described
+in the next section of this paper.
 
@@
   \captionof{figure}{a) box plots of 25 crowd-sourced sets of fission
-    track counts for 70 analysts. b) radial plot comparing the paired
-    track counts for the first two grains in the dataset (4649 and
-    4673). The estimate of user PV is shown in blue. Analyst AC used a
-    different ROI and is not shown.\medskip}
+    track counts for first-time student 70 analysts using ROIs defined
+    by PV. b) radial plot comparing the paired track counts for the
+    first two grains in the dataset (4649 and 4673). The estimate of
+    user PV is shown in blue. Analyst AC used a different ROI and is
+    not shown.\medskip}
   \label{fig:radialcrowd}
@@
 by multiple analysts is prohibitively expensive in a normal laboratory
-environment. geochron@home changes this, by bringing fission
-track dating to a proverbial `county fair' of citizen
-scientists.\medskip
+environment. geochron@home changes this, by bringing fission track
+analysis to a proverbial `county fair' of citizen scientists.\medskip
 
-Figure~\ref{fig:crowdsourcing} presents the results of a
-crowd-sourcing experiment that was carried out at UCL as part of an
-undergraduate course in isotope geology. 70~students were asked to
-create an account on the geochron@home server by selecting a
-unique username and password. After spending a few minutes to complete
-the compulsory tutorial (Section~\ref{sec:tutorial}), the students
-were asked to count fission tracks in randomly assigned image stacks
-of Mount Dromedary apatite. Each student was required to analyse at
-least 15 grains, using pre-defined ROIs. In a matter of hours, the
-students amassed a dataset of 37,331 fission track counts in 34
-grains. This large dataset was parsed into separate data files --one
-for each student-- which they processed during an assessed practical
+Figure~\ref{fig:radialcrowd} presents the results of a crowd-sourcing
+experiment that was carried out at UCL as part of an undergraduate
+course in isotope geology. 70~students were asked to create an account
+on the geochron@home server by selecting a unique username and
+password. After spending a few minutes to complete the compulsory
+tutorial (Section~\ref{sec:tutorial}), the students were asked to
+count fission tracks in randomly assigned image stacks of Mount
+Dromedary apatite. Each student was required to analyse at least 15
+grains, using pre-defined ROIs. In a matter of hours, the students
+amassed a dataset of 37,331 fission track counts in 34 grains. This
+large dataset was parsed into separate data files --one for each
+student-- which they processed during an assessed practical
 exercise.\medskip
@@
 The raw counts are shown as a series of box plots in
-Figure~\ref{fig:crowdsourcing}a for the same selection of 25 grains
-that were analysed by AC and PV in Section~\ref{sec:private}. The box
-plots exhibit significant variability between the grains and between
-the students. This inter-grain variability is expected. It reflects
-the different sizes of the ROIs, and the variable U-content of the
-grains. The variability of the track counts for each grain results
-from observer bias. This phenomenon was previously discussed in
+Figure~\ref{fig:radialcrowd}a for the same selection of 25 grains that
+were analysed by AC and PV in Section~\ref{sec:private}. The box plots
+exhibit significant variability between the grains and between the
+students. The inter-grain variability is expected. It reflects the
+different sizes of the ROIs, and the variable U-content of the
+grains. The inter-student variability is caused by observer bias. This
+phenomenon was previously discussed in
 Section~\ref{sec:private}. However, the bias is much stronger for the
 crowd-sourced dataset than it was for the two `expert' analysts of
-Section~\ref{sec:private}.\medskip
+Section~\ref{sec:private}. Figure~\ref{fig:radialcrowd}b compares the
+42 raw fission track counts of the first two grains on a radial plot,
+where
+\begin{equation}
+  \frac{\mbox{se}(N_1/N_2)}{N_1/N_2}
+  \approx
+  \sqrt{
+    \frac{1}{N_1} + \frac{1}{N_2}
+  }
+  \label{eq:N12}
+\end{equation}
 
-Figure~\ref{fig:crowdsourcing}b compares the 42 raw fission track
-counts of the first two grains on a radial plot.  The pairwise ratios
-of the counts for each analyst are more consistent than the raw counts
-of Figure~\ref{fig:crowdsourcing}a.  Students who over- or undercount
-the first grain, tend to over- or undercount the second grain as
-well. Nevertheless, the results are overdispersed with respect to the
-counting uncertainties, even when using the `pessimistic' scenario of
-Equation~\ref{eq:pessimistic}.  The poor consistency of the results
-reflects the very limited training received by the students. But
-despite these inconsistencies, the opinion of an `expert' (PV, shown
-in blue) falls the middle of the estimates obtained by the citizen
-scientists. This suggests that the collective wisdom of the entire
-group of students is greater than that of any individual
-analyst.\medskip
+The pairwise ratios of the counts for each analyst are more consistent
+than the raw counts of Figure~\ref{fig:crowdsourcing}a.  Students who
+over- or undercount the first grain, tend to over- or undercount the
+second grain as well. Nevertheless, the results are overdispersed with
+respect to the counting uncertainties. The overdispersion is
+particularly bad considering that Equation~\ref{eq:N12} ignores the
+systematic uncertainty caused by the fact that all analysts counted
+exactly the same grains.  The poor consistency of the results reflects
+the very limited training received by the students. But despite these
+inconsistencies, the opinion of an `expert' (PV, shown in blue) falls
+the middle of the estimates obtained by the citizen scientists. This
+suggests that the collective wisdom of the entire group of students is
+greater than that of any individual analyst.\medskip
 
-As metioned in Section~\ref{sec:architecture}, geochron@home stores
+As mentioned in Section~\ref{sec:architecture}, geochron@home stores
 the actual track positions marked by the users. This raw data can be
 downloaded as a \texttt{.json} file and inspected in detail.
-Figure~\ref{fig:radialcrowd} is a two-dimensional histogram for the
+Figure~\ref{fig:crowdsourcing} is a two-dimensional histogram for the
 x-y positions of all the track positions generated by the students in
@@
 
-Figure~\ref{fig:radialcrowd} also shows that some students counted the
-tails of the fission tracks rather than their etch pits, despite being
-told the opposite in the tutorial. Fixing this issue will require some
-improvements to the tutorial pages (Section~\ref{sec:outlook}).
+Figure~\ref{fig:crowdsourcing} also shows that some students counted
+the tails of the fission tracks rather than their etch pits, despite
+being told the opposite in the tutorial. Fixing this issue will
+require some improvements to the tutorial pages
+(Section~\ref{sec:outlook}).
 
@@
   fission tracks automatically. Following the model of Fission Track
-  Studio, machine learning can be used to reverse the fission track
-  counting process. Instead of asking users to count the fission
-  tracks in a sample, the software can ask them to check the results
-  proposed by an AI algorithm, and to remove any features that are
-  \emph{not} fission tracks.\medskip
+  Studio \citep{gleadow2009, gleadow2019}, machine learning can be
+  used to reverse the fission track counting process. Instead of
+  asking users to count the fission tracks in a sample, the software
+  can ask them to check the results proposed by an AI algorithm, and
+  to remove any features that are \emph{not} fission tracks.\medskip
 
@@
 These four improvements will be made by ourselves pending additional
-funding. However, because geochron@home is free and open
-software, we invite any interested parties to join the effort and
-extend or improve our code.
+funding. However, because geochron@home is free and open software, we
+invite any interested parties to join the effort and extend or improve
+our code.
 
@@
 
-Another advantage of the geochron@home wokflow is the
-separation of image acquisition and image analysis. This separation
-reduces the hardware requirements for fission track dating. It opens
-up the possibility to share resources. State-of-the-art digital
-microscopes are expensive. Using geochron@home, a single
-microscope can serve multiple users and make fission track analysis
-more affordable.\medskip
+Another advantage of the geochron@home wokflow is the separation of
+image acquisition and image analysis. This separation reduces the
+hardware requirements for fission track geochronology. It opens up the
+possibility to share resources. State-of-the-art digital microscopes
+are expensive. Using geochron@home, a single microscope can serve
+multiple users and make fission track analysis more
+affordable.\medskip
 
@@
 and tools such as overdispersion, mixture modelling and radial plots,
-which have been commonplace in fission track dating for decades
+which have been commonplace in fission track analysis for decades
 \citep{vermeesch2019b}. In a similar vein, the subjective nature of
@@
 
-With the development of geochron@home, fission tracks is once
-again ahead of the pack in terms of FAIR data analysis.
-geochron@home currently only stores images and counts. This
-is enough to reproduce the results of fission track studies using the
-external detector method, but not for LA-ICP-MS based data. FAIR data
+With the development of geochron@home, fission track geochronology is
+once again ahead of the pack in terms of FAIR data analysis.
+geochron@home currently only stores images and counts. This is enough
+to reproduce the results of fission track studies using the external
+detector method, but not for LA-ICP-MS based data. FAIR data
 processing of LA-ICP-MS data requires a new generation of mass
-spectrometer data reduction software. We are currently working on
-this. The development of FAIR ICP-MS data pipelines will not only
-benefit fission track dating but other chronometers as well, such as
-in-situ U--Pb, Rb--Sr and Lu--Hf \citep{vermeesch2025}.\medskip
+spectrometer data reduction software. We are currently working on this
+\citep{vermeesch2025b}. The development of FAIR ICP-MS data pipelines
+will not only benefit fission track analysis but other chronometers as
+well, such as in-situ U--Pb, Rb--Sr and Lu--Hf.\medskip
 
-With the establishment of FAIR data, the era of `fiction tracks' will
-be gone for good, and geochronology will be well placed to avoid the
-reproducibility problems that have plagued other fields of science.
+With the establishment of FAIR data, geochronology will be well placed
+to avoid the reproducibility problems that have plagued other fields
+of science.
 
@@
   provided the samples, prepared the training data and counted fission
-  tracks. Rex Galbraith derived Equation~\ref{eq:optimistic}.}
+  tracks. Rex Galbraith derived Equation~\ref{eq:NO}.}
 

### 2025-05-09 geochron@home/R/N00.R

@@ -1,40 +1,0 @@
-# to be merged with helper.R and/or main.R
-
-rm(list=ls())
-graphics.off()
-
-source("helper.R")
-source("polygon_overlap.R")
-
-json <- IsoplotR:::fromJSON(file='../json/results.json')
-roiss <- IsoplotR:::fromJSON(file='../json/roiss.json')
-Pieter <- IsoplotR:::fromJSON(file='../json/Pieter.json')
-Andy <- IsoplotR:::fromJSON(file='../json/Andy.json')
-
-PieterResults <- parseJSON(Pieter,defaultrois=roiss)
-AndyResults <- parseJSON(Andy)
-grains <- intersect(names(PieterResults),names(AndyResults))
-
-count <- function(xyP,xyA,i=0,cutoff=20){
-    nP <- nrow(xyP)
-    nA <- nrow(xyA)
-    xy <- rbind(xyP,xyA)
-    d <- dist(xy)
-    dPA <- as.matrix(d)[1:nP,(nP+1):(nP+nA)]
-    j <- which.min(dPA)
-    if (dPA[j]<cutoff & nP>1 & nA>1){
-        rP <- row(dPA)[j]
-        rA <- col(dPA)[j]
-        i <- count(xyP=xyP[-rP,,drop=FALSE],
-                   xyA=xyA[-rA,,drop=FALSE],
-                   i=i+1,cutoff=cutoff)
-    }
-    i
-}
-
-N00 <- c()
-for (grain in grains){
-    xyP <- PieterResults[[grain]]$counts
-    xyA <- AndyResults[[grain]]$counts
-    N00[grain] <- count(xyP,xyA)
-}

### 2025-05-09 geochron@home/R/helper.R

@@ -172,7 +172,8 @@ list2table <- function(lst){
     data.frame(index=index,grain=grains,worker=workers,Ns=Ns,A=A)
 }
 
-count <- function(xyP,xyA,i=0,cutoff=20){
+# counts tracks identified by analysts P and A within a cutoff distance
+getN00 <- function(xyP,xyA,i=0,cutoff=20){
     nP <- nrow(xyP)
     nA <- nrow(xyA)
     xy <- rbind(xyP,xyA)
@@ -182,9 +183,11 @@ count <- function(xyP,xyA,i=0,cutoff=20){
     if (dPA[j]<cutoff & nP>1 & nA>1){
         rP <- row(dPA)[j]
         rA <- col(dPA)[j]
-        i <- count(xyP=xyP[-rP,,drop=FALSE],
-                   xyA=xyA[-rA,,drop=FALSE],
-                   i=i+1,cutoff=cutoff)
+        iPA <- getN00(xyP=xyP[-rP,,drop=FALSE],
+                      xyA=xyA[-rA,,drop=FALSE],
+                      i=i+1,cutoff=cutoff)
+    } else {
+        iPA <- c(N00=i,nP=nP,nA=nA)
     }
-    i
+    return(iPA)
 }

### 2025-05-09 geochron@home/R/main.R

@@ -1,6 +1,8 @@
 rm(list=ls())
 graphics.off()
 
+library(IsoplotR)
+library(provenance)
 library(ggplot2)
 library(jpeg)
 library(grid)
@@ -19,33 +21,34 @@ AndyResults <- parseJSON(Andy)
 grains <- intersect(names(PieterResults),names(AndyResults))
 
 #### 1. calculate the density ratios for PV and AC ####
-PAsPA1 <- PAsPA2 <- PAsPA3 <- NULL
+PAsPA <- NULL
+total_counts_in_A0 <- c(N00=0,nP=0,nA=0)
+total_counts <- c(N1=0,N2=0)
+total_area <- c(A0=0,A1=0,A2=0)
 for (i in seq_along(grains)){
     grain <- grains[i]
-    PieterNsA <- grain2NsA(PieterResults[[grain]])
-    AndyNsA <- grain2NsA(AndyResults[[grain]])
-    A0 <- polygon_overlap_area(PieterResults[[grain]]$ROI,
-                               AndyResults[[grain]]$ROI)
-    # optimistic scenario: Rex Galbraith formula
-    PA <- (PieterNsA[1]*AndyNsA[2])/(PieterNsA[2]*AndyNsA[1])
-    # pessimistic scenario: assuming independence
-    N1 <- PieterNsA[1]
-    N2 <- AndyNsA[1]
-    sPA1 <- PA * sqrt(1/N1 + 1/N2)
-    PAsPA1 <- rbind(PAsPA1,c(PA,sPA1))
-    rho <- (N1+N2)/(PieterNsA[2]+AndyNsA[2])
-    sPA2 <- PA * sqrt(1/(rho*PieterNsA[2]) +
-                      1/(rho*AndyNsA[2]) -
-                      2*A0/(rho*PieterNsA[2]*AndyNsA[2]))
-    PAsPA2 <- rbind(PAsPA2,c(PA,sPA2))
-    # realistic scenario: Rex Galbraith formula 2
+    ROIP <- PieterResults[[grain]]$ROI
+    ROIA <- AndyResults[[grain]]$ROI
+    ROI0 <- polygon_intersection(ROIP,ROIA)
+    A1 <- polygon_area(ROIP)
+    A2 <- polygon_area(ROIA)
+    A0 <- polygon_area(ROI0)
[…]

### 2025-05-09 geochron@home/R/polygon_overlap.R

@@ -5,13 +5,18 @@ polygon_area <- function(vertices) {
   n <- nrow(vertices)
   area <- 0
   for (i in 1:(n-1)) {
-    area <- area + (vertices[i,1] * vertices[i+1,2]) - (vertices[i+1,1] * vertices[i,2])
+      area <- area +
+          (vertices[i,1] * vertices[i+1,2]) -
+          (vertices[i+1,1] * vertices[i,2])
   }
-  area <- area + (vertices[n,1] * vertices[1,2]) - (vertices[1,1] * vertices[n,2])
+  area <- area +
+      (vertices[n,1] * vertices[1,2]) -
+      (vertices[1,1] * vertices[n,2])
   return(abs(area) / 2)
 }
 
-# Function to check if a point is inside a polygon using the ray-casting algorithm
+# Function to check if a point is inside
+# a polygon using the ray-casting algorithm
 point_in_polygon <- function(point, polygon) {
   n <- nrow(polygon)
   x <- point[1]
@@ -104,7 +109,8 @@ polygon_intersection <- function(poly1, poly2) {
     
     # Sort points counterclockwise using centroid
     centroid <- colMeans(intersection_points)
-    angles <- atan2(intersection_points[,2] - centroid[2], intersection_points[,1] - centroid[1])
+    angles <- atan2(intersection_points[,2] - centroid[2],
+                    intersection_points[,1] - centroid[1])
     intersection_points <- intersection_points[order(angles), ]
     
     return(intersection_points)
@@ -121,4 +127,14 @@ polygon_overlap_area <- function(poly1, poly2) {
   } else {
     return(0)
   }
+}
+
+polygon_test <- function(poly1,poly2){
+    poly3 <- polygon_intersection(poly1,poly2)
+    xlim <- range(c(poly1[,1],poly2[,1]))
+    ylim <- range(c(poly1[,2],poly2[,2]))
+    plot(xlim,ylim,type='n')
+    polygon(poly1,col='red')
+    polygon(poly2,col='blue')
+    polygon(poly3,col='green')
 }

### 2025-07-02 geochron@home/manuscript/paper.tex
Minor edit to manuscript
geochron@home

@@
 combine fission track counts for multiple users within a lab group;
-(2) the use of the geochron@home archive in the peer review system;
-(3) the effectiveness of even simple tutorials in teaching novice
-users how to count fission tracks; and (4) the `wisdom of crowds' in
-fission track identification.\medskip
+(2) the value of the geochron@home archive in the peer review system;
+(3) the use of simple tutorials in teaching novice users how to count
+fission tracks; and (4) the `wisdom of crowds' in fission track
+identification.\medskip
 
@@
 \item Upload the data to the geochron@home platform. geochron@home is
-  a Django web-app featuring a user interface powered by the Leaflet
-  library (JavaScript).  The database can be accessed via a Python API
-  and a (more limited) web-based GUI. Accessing the API requires
-  administrator privileges. Administrators can create projects,
-  samples and grains; download data; and set the access rights of
-  `ordinary' users.  Projects can be private or public. The source
-  code and installation instructions for geochron@home are freely
-  available over GitHub (see the Data Availability statement at the
-  end of this paper). This allows fission track users to set up their
-  own server. Alternatively, fission track labs can upload their data
-  to the UCL server by contacting the corresponding author.
+  a Django web-app with a PostgreSQL database.  The database can be
+  accessed via a Python API and a (more limited) web-based
+  GUI. Accessing the API requires administrator
+  privileges. Administrators can create projects, samples and grains;
+  download data; and set the access rights of `ordinary' users.
+  Projects can be private or public. The source code and installation
+  instructions for geochron@home are freely available over GitHub (see
+  the Data Availability statement at the end of this paper). This
+  allows fission track users to set up their own
+  server. Alternatively, fission track labs can upload their data to
+  the UCL server by contacting the corresponding author.
 
 \item Analyse the images with a browser-based `virtual fission track
-  microscope'. This virtual instrument has a simple user interface
-  with controls to zoom, pan and focus in or out of the digital image
-  stack. Depending on the permissions granted to the user by the
-  administrator, the virtual microsocope offers a number of different
-  options.  Entry level `ordinary' users are only allowed to count
-  tracks by clicking within a pre-defined `region of interest'. In
-  contrast, `superusers' are allowed to define their own regions of
-  interest. Once the user is satisfied that they have counted all the
-  fission tracks in a particular grain, they can submit the results to
-  the server. They are then presented with a new set of images until
-  all grains are counted.
+  microscope' powered by the Leaflet library (JavaScript). This
+  virtual instrument acts as a front-end to geochron@home. It has a
+  simple user interface with controls to zoom, pan and focus in or out
+  of the digital image stack. Depending on the permissions granted to
+  the user by the administrator, the virtual microsocope offers a
+  number of different options.  Entry level `ordinary' users are only
+  allowed to count tracks by clicking within a pre-defined `region of
+  interest'. In contrast, `superusers' are allowed to define their own
+  regions of interest. Once the user is satisfied that they have
+  counted all the fission tracks in a particular grain, they can
+  submit the results to the server. They are then presented with a new
+  set of images until all grains are counted.
 
@@
 from experienced analysts and future plans include the addition of
-fully interactive `quizzes' (Section~\ref{sec:conclusions}). Despite
-the simplicity of the current tutorials, they are already quite
-effective, as demonstrated by the crowd-sourcing experiment described
-in the next section of this paper.
+fully interactive `quizzes' (Section~\ref{sec:conclusions}). The
+limitations of the current tutorials are apparent in the results of
+the crowd-sourcing experiment described in the next section of this
+paper.
 
@@
 %\begin{figure}[!ht]
-{ \centering \includegraphics[width=15cm]{radialcrowd.pdf}
+{ \centering \includegraphics[width=\linewidth]{radialcrowd.pdf}
   \captionof{figure}{a) box plots of 25 crowd-sourced sets of fission
-    track counts for first-time student 70 analysts using ROIs defined
-    by PV. b) radial plot comparing the paired track counts for the
-    first two grains in the dataset (4649 and 4673). The estimate of
-    user PV is shown in blue. Analyst AC used a different ROI and is
-    not shown.\medskip}
+    track counts for 69 analysts; b) comparison of grains~23 and 25,
+    which were chosen because of their high counts and variability; c)
+    radial plot of the pairwise ratios of the fission track counts for
+    grains~23 and 25. The results of analyst PV are shown in
+    blue.\medskip}
   \label{fig:radialcrowd}
@@
 experiment that was carried out at UCL as part of an undergraduate
-course in isotope geology. 70~students were asked to create an account
+course in isotope geology. 68~students were asked to create an account
 on the geochron@home server by selecting a unique username and
@@
 crowd-sourced dataset than it was for the two `expert' analysts of
-Section~\ref{sec:private}. Figure~\ref{fig:radialcrowd}b compares the
-42 raw fission track counts of the first two grains on a radial plot,
-where
+Section~\ref{sec:private}. The majority of the student analysts
+undercounted the fission tracks by up to a factor of five compared to
+the `expert' analyst (PV, shown in blue). The poor consistency of the
+results reflects the very limited training received by the
+students.\medskip
+
+Grains 23 and 25 have greater than average track densities, and are
+particularly dispersed. Plotting the 37 paired counts for each user on
+a scatter plot reveals a strong correlation
+(Figure~\ref{fig:radialcrowd}b). This means that students who over- or
+undercount the first grain, also tend to over- or undercount the
+second grain. Even though most students count fewer fission tracks
+than PV, the ratio of their counts are similar.
+Figure~\ref{fig:radialcrowd}c compares grains 23 and 25 on a radial
+plot, where
 \begin{equation}
@@
 
-The pairwise ratios of the counts for each analyst are more consistent
-than the raw counts of Figure~\ref{fig:crowdsourcing}a.  Students who
-over- or undercount the first grain, tend to over- or undercount the
-second grain as well. Nevertheless, the results are overdispersed with
-respect to the counting uncertainties. The overdispersion is
-particularly bad considering that Equation~\ref{eq:N12} ignores the
-systematic uncertainty caused by the fact that all analysts counted
-exactly the same grains.  The poor consistency of the results reflects
-the very limited training received by the students. But despite these
-inconsistencies, the opinion of an `expert' (PV, shown in blue) falls
-the middle of the estimates obtained by the citizen scientists. This
-suggests that the collective wisdom of the entire group of students is
-greater than that of any individual analyst.\medskip
+The tight clustering of the data within the 2-se band confirms that
+the pairwise ratios of the counts for each analyst are more consistent
+than the raw counts of Figure~\ref{fig:crowdsourcing}a.  Despite the
+tremendous scatter of the raw track counts obtained by the citizen
+scientists, their pairwise ratio estimates scatter evenly around the
+expert opinion. This suggests that the collective wisdom of the entire
+group of students is greater than that of any individual
+analyst.\medskip
 

### 2025-07-02 geochron@home/R/helper.R

@@ -191,3 +191,66 @@ getN00 <- function(xyP,xyA,i=0,cutoff=20){
     }
     return(iPA)
 }
+
+clean_results <- function(results,ACgrains){
+    time_obj <- sapply(results$create_date,
+                       FUN=as.POSIXct,
+                       format = "%Y-%m-%d %H:%M:%OS",
+                       tz = "UTC")
+    timely_submission <- time_obj < as.POSIXct("2025-02-11 00:00:00.00+00:00",
+                                               format = "%Y-%m-%d %H:%M:%OS", tz = "UTC")
+    is_trustworthy <- !(results$user_id %in% c(237,238)) # non UCL students
+    uses_default_roi <- !(results$user_id %in% c(292,2)) # AC and PV
+    is_admin <- results$user_id == 1
+    AC_ids <- unique(results$index[results$user_id==292])
+    good_grains <- results$index %in% AC_ids
+    good_counts <- (timely_submission & is_trustworthy & uses_default_roi) | is_admin
+    keep <- good_grains & good_counts
+    results[keep,]
+}
+
+add_admin_count_to_boxplot <- function(trustworthy_results,grouped_list){
+    admin_results <- trustworthy_results[trustworthy_results$user_id == 1,]
+    ACids <- as.numeric(names(grouped_list))
+    admin_counts <- admin_results$count[match(ACids,admin_results$index)]
+    points(admin_counts,1:length(ACids),pch=21,bg='blue')
+}
+
+compare_grains <- function(results,grain1,grain2,plot=TRUE,...){
+    results1 <- results[results$index == grain1,]
+    results2 <- results[results$index == grain2,]
+    users <- intersect(results1$user_id,
+                       results2$user_id)
+    i1 <- match(users,results1$user_id)
+    i2 <- match(users,results2$user_id)
+    out <- list(name=paste0(grain1,' vs ',grain2))
+    class(out) <- 'counts'
+    out$x <- cbind(results1$count[i1],
+                   results2$count[i2])
+    colnames(out$x) <- c(paste0('Ns(',grain1,')'),
+                         paste0('Ns(',grain2,')'))
+    ns <- nrow(out$x)
+    if (plot){
+        bg <- rep(NA,ns)
+        iadmin <- which(users==1)
+        bg[iadmin] <- 'blue'
+        plot(out$x,
[…]

### 2025-07-02 geochron@home/R/main.R

@@ -11,10 +11,11 @@ library(gridExtra)
 source("helper.R")
 source("polygon_overlap.R")
 
-json <- IsoplotR:::fromJSON(file='../json/results.json')
-roiss <- IsoplotR:::fromJSON(file='../json/roiss.json')
-Pieter <- IsoplotR:::fromJSON(file='../json/Pieter.json')
-Andy <- IsoplotR:::fromJSON(file='../json/Andy.json')
+results <- read.csv('../data/results.csv')
+json <- IsoplotR:::fromJSON(file='../data/results.json')
+roiss <- IsoplotR:::fromJSON(file='../data/roiss.json')
+Pieter <- IsoplotR:::fromJSON(file='../data/Pieter.json')
+Andy <- IsoplotR:::fromJSON(file='../data/Andy.json')
 
 PieterResults <- parseJSON(Pieter,defaultrois=roiss)
 AndyResults <- parseJSON(Andy)
@@ -121,33 +122,31 @@ grid.arrange(
 par(op)
 dev.off()
 
-#### 6. Create a boxplot of all the crowdsourcing results ####
-####    and a radial plot of the two grains in part 5.    ####
-pdf(file='../output/radialcrowd.pdf',width=10,height=5)
-op <- par(mfrow=c(1,2))
-tab <- list2table(allgrains[grains])
-# remove Andy Carter, who used a different count area:
-dat <- tab[tab[,'worker'] != 292,]
-ip <- par(mar=c(4,4,1,1))
-boxplot(Ns ~ index,data=dat,col=NA)
-legend('topleft',legend="a)",bty='n')
-par(ip)
-hasgrain1 <- which(dat[,'grain'] == grain1)
-hasgrain2 <- which(dat[,'grain'] == grain2)
-userswithgrains12 <- intersect(dat[hasgrain1,'worker'],dat[hasgrain2,'worker'])
-inum <- which(dat[hasgrain1,'worker'] %in%  userswithgrains12)
-iden <- which(dat[hasgrain2,'worker'] %in%  userswithgrains12)
-num <- dat[hasgrain1[inum],'Ns']
-den <- dat[hasgrain2[iden],'Ns']
-numden <- cbind(num,den)
-index1 <- which(unique(dat[,'grain']) %in% grain1)
-index2 <- which(unique(dat[,'grain']) %in% grain2)
-colnames(numden) <- c(paste0('Ns(',index1,')'),
[…]

### 2025-07-21 carbonatePbUproblems/Fig6.R
Added tables to broken chronometer repo

@@ -30,7 +30,7 @@ text(tt[last]-0.1,terr[last]+5,label,pos=3,xpd=NA)
 
 # panel b)
 
-maxerr <- 140 # y axis limit
+maxerr <- 500 # y axis limit
 plot(range(tt),c(1,maxerr),type='n',xlab=expression("t"[c]*"(Ma)"),
      ylab=expression("Uncertainty of the correction ([t"[u]*"-t"[l]*"]/t"[c]*", %)"),bty='n',log='y')
 legend('topright','b)',bty='n')
@@ -63,7 +63,6 @@ for (i in seq_along(U48i)){
         label <- U48i[i]
     }
     text(tt[last],terr[last],label,pos=2)
-
 }
 
 # worst case scenario envelope

### 2025-07-21 carbonatePbUproblems/main.R

@@ -1,8 +1,10 @@
 library(IsoplotR)
 
-source("plot.R")
+source('plot.R')
 
-source("MonteCarlo.R")
+source('MonteCarlo.R')
+
+source('tables.R')
 
 source('Fig1.R')
 

### 2025-07-21 carbonatePbUproblems/tables.R

@@ -1,0 +1,22 @@
+d <- diseq(U48=list(x=1.006,option=2,M=100),
+           PaU=list(x=0,option=1))
+ttrue <- seq(from=0.5,to=3,by=0.5)
+t75 <- t68 <- i48 <- NA*ttrue
+for (i in seq_along(ttrue)){
+    McL <- IsoplotR:::mclean(tt=ttrue[i],d=d)
+    i48[i] <- d1$U48$x <- McL$U48i
+    t68[i] <- age(x=McL$Pb206U238,method='U238-Pb206')[1]
+    t75[i] <- age(x=McL$Pb207U235,method='U235-Pb207')[1]
+}
+
+print(
+    rbind(
+        t = ttrue,
+        maxU48i = i48,
+        relerr = 100*(t68-ttrue)/ttrue)
+)
+
+print(
+    rbind(t = ttrue,
+          relerr = 100*(ttrue-t75)/ttrue)
+)

### 2025-07-21 geochron@home/manuscript/paper.tex
Started implementing Rex Galbraith suggestions

@@
 
-The two analysts defined their own ROIs, resulting in slightly
-different density estimates (Figure~\ref{fig:AvP}a) for 25 apatite
-grains.  Let $N_{PV}$ and $N_{AC}$ be the number of spontaneous tracks
-counted by PV and AC over areas $A_{PV}$ and $A_{AC}$,
-respectively. Then the ratio of the estimated track densities are
-given by $\hat{\rho}_{PV} = N_{PV}/A_{PV}$ and $\hat{\rho}_{AC} =
-N_{AC}/A_{AC}$. The uncertainties of these density ratios are strongly
-correlated due to the significant overlap between the counting areas
-chosen by the two analysts:
+Let $N_1$ and $N_2$ be the numbers of spontaneous tracks counted by PV
+and AC over areas $A_1$ and $A_2$, respectively. Then their estimated
+track densities are give by $\hat{\rho}_1 = N_1/A_1$ and $\hat{\rho}_2
+= N_2/A_2$. In this situation when the two areas overlap, $N_1$ and
+$N_2$ cannot be treated as independent Poisson counts because the two
+analysts will count some of the same tracks. It can be shown that,
+under simple (ideal) assumptions, the uncertainty of the ratio of the
+estimated track densities is given approximately by
 \begin{equation}
-  \frac{\mbox{se}(\hat{\rho}_{PV}/\hat{\rho}_{AC})}{\hat{\rho}_{PV}/\hat{\rho}_{AC}}
-  \approx
-  \sqrt{
-    \frac{1}{N_{PV}} + \frac{1}{N_{AC}} - \frac{2\,N_{0}}{N_{PV}\,N_{AC}}
-  }
+  \frac{\mbox{se}(\hat{\rho}_1/\hat{\rho}_2)}{\hat{\rho}_1/\hat{\rho}_2} \approx
+  \sqrt{ \frac{1}{N_1} + \frac{1}{N_2} - \frac{2\,N_0}{N_1\,N_2} }
   \label{eq:NO}
@@
 
-\noindent where $N_{0}$ is the number of tracks counted by both
+\noindent where $N_0$ is the number of tracks counted by both
 observers in the area of overlap ($A_0$, say) between their respective
-ROIs.  If there is no overlap between the two counting areas (i.e., if
-$A_0=0$), then Equation~\ref{eq:NO} reduces to the conventional
-coefficient of variation for the ratio of two independent Poisson
-variables \citep{galbraith1990a}.  Conversely, if the counting areas
-are identical (i.e., if $A_{PV}=A_{AC}=A_0$), and if the two analysts
-produce perfectly consistent counts (i.e. if $N_{PV} = N_{AC} =
-N_{0}$), then Equation~\ref{eq:NO} returns a zero value.\medskip
+ROIs.\medskip
 
 The radial plot of Figure~\ref{fig:AvP}b contains two sets of counts,
-with $\sum N_{PV} = 686$ and $\sum N_{AC} = 679$, and a weighted mean
-track density ratio $\hat{\rho}_{PV}/\hat{\rho}_{PV} = 0.94$ with
+with $\sum N_{1} = 686$ and $\sum N_{2} = 679$, and a weighted mean
+track density ratio $\hat{\rho}_{1}/\hat{\rho}_{1} = 0.94$ with
 relative standard error 0.013. PV counted 600 tracks in those common
@@
 %\begin{figure}[!ht]
-{ \centering \includegraphics[width=12cm]{AvP.pdf}
+{ \centering \includegraphics[width=\linewidth]{AvP.pdf}
   \captionof{figure}{Comparison of fission track data for Mount
@@
     Andrew Carter). a) the track counts and ROIs of both users for
-    grain 4648, shown in blue (PV) and red (AC); b) radial plot for
-    the ratio of the track densities of both users for all 25 grains
-    analysed by the two analysts, using Equation~\ref{eq:NO}.
-    \medskip}
+    grain 4648, shown in blue (PV) and red (AC); b) comparison of the
+    track densities for PV and AC for all 25 grains analysed by the
+    two analysts with a reference line for
+    $\hat{\rho}_\text{PV}=\hat{\rho}_\text{AC}$; c) radial plot of the
+    same data, using Equation~\ref{eq:NO}.}
   \label{fig:AvP}
@@
 
-Figure~\ref{fig:radialcrowd} presents the results of a crowd-sourcing
-experiment that was carried out at UCL as part of an undergraduate
-course in isotope geology. 68~students were asked to create an account
-on the geochron@home server by selecting a unique username and
-password. After spending a few minutes to complete the compulsory
-tutorial (Section~\ref{sec:tutorial}), the students were asked to
-count fission tracks in randomly assigned image stacks of Mount
-Dromedary apatite. Each student was required to analyse at least 15
-grains, using pre-defined ROIs. In a matter of hours, the students
-amassed a dataset of 37,331 fission track counts in 34 grains. This
-large dataset was parsed into separate data files --one for each
-student-- which they processed during an assessed practical
-exercise.\medskip
+Figure~\ref{fig:radialcrowd}a shows box plots and summary statistics
+for the raw counts made by 68 students (two of the students were
+unable to complete the assignment) for the same selection of 25 grains
+that were analysed by AC and PV in Section~\ref{sec:private}. Everyone
+counted tracks in the same ROIs, and counts made by PV for those ROIs
+have been added as blue dots.  Variation between the mean values is to
+be expected because of the differing numbers of tracks due to varying
+areas and uranium contents of the grains. But the variation between
+counts within grains (shown by the standard deviations) is due
+entirely to differences between the students' recognising and counting
+exactly the same tracks. These standard deviations increase with the
+mean number of tracks and are considerable in size, being on average
+about 40\% of the mean.  Also, the vast majority of students counted
+fewer tracks than PV did, often many fewer. In 21 of the 25 grains,
+PV's count is above the upper quartile of the students', and on 2
+occasions someone counted no tracks at all.  With expert trained
+counters one would expect much smaller differences between counts.
+Nevertheless there are many repeated counts also.\medskip
 
-The raw counts are shown as a series of box plots in
-Figure~\ref{fig:radialcrowd}a for the same selection of 25 grains that
-were analysed by AC and PV in Section~\ref{sec:private}. The box plots
-exhibit significant variability between the grains and between the
-students. The inter-grain variability is expected. It reflects the
-different sizes of the ROIs, and the variable U-content of the
-grains. The inter-student variability is caused by observer bias. This
-phenomenon was previously discussed in
-Section~\ref{sec:private}. However, the bias is much stronger for the
-crowd-sourced dataset than it was for the two `expert' analysts of
-Section~\ref{sec:private}. The majority of the student analysts
-undercounted the fission tracks by up to a factor of five compared to
-the `expert' analyst (PV, shown in blue). The poor consistency of the
-results reflects the very limited training received by the
-students.\medskip
-
-Grains 23 and 25 have greater than average track densities, and are
-particularly dispersed. Plotting the 37 paired counts for each user on
-a scatter plot reveals a strong correlation
-(Figure~\ref{fig:radialcrowd}b). This means that students who over- or
-undercount the first grain, also tend to over- or undercount the
-second grain. Even though most students count fewer fission tracks
-than PV, the ratio of their counts are similar.
-Figure~\ref{fig:radialcrowd}c compares grains 23 and 25 on a radial
-plot, where
+Figure~\ref{fig:radialcrowd}b shows a scatter plot of pairs of counts
+for two grains made by 33 students who counted both of them. They
+average around 23 and 30 tracks per grain compared with PV's pair of
+44 and 63, which is highlighted in blue. There is a positive
+correlation between the pairs, consistent with systematic observer
+effects (i.e., people who count a low value in one grain tend to count
+similarly low in the other, and vice versa).  Comparing other pairs of
+grains shows similar results, with correlations varying between 0.3
+and 0.9. In all cases we have looked at, the residual variation, after
+allowing for systematic differences between students, is consistent
+with Poisson variation.  Figure~\ref{fig:radialcrowd}c shows a radial
+plot of the ratios of the counts in Panel b, using the precisions
+given by
 \begin{equation}
@@
 
-The tight clustering of the data within the 2-se band confirms that
-the pairwise ratios of the counts for each analyst are more consistent
-than the raw counts of Figure~\ref{fig:crowdsourcing}a.  Despite the
-tremendous scatter of the raw track counts obtained by the citizen
-scientists, their pairwise ratio estimates scatter evenly around the
-expert opinion. This suggests that the collective wisdom of the entire
-group of students is greater than that of any individual
-analyst.\medskip
+The context here is rather different from that in
+Figure~\ref{fig:AvP}b, where we were comparing the ratios of counts
+made by 2 analysts on each of 25 grains. Here we are comparing the
+ratios of counts for the same 2 grains made by each of 33
+analysts. There is no theoretical justification for
+Equation~\ref{eq:N12} here, but it still provides a useful benchmark
+for assessing variation between the ratios. In
+Figure~\ref{fig:radialcrowd}c, that variation is consistent with
+Poisson variation (as it is for most other pairs of grains) but with
+expert counters one would expect less variation than that. However, in
+spite of the under-counting by the students, the pooled the students'
+ratio of 0.76 is quite close to PV's ratio of 0.70. This happens quite
+often in this data set, but not always, and there are some pairs of
+grains where the students' pooled ratio differs somewhat from
+PV's.\medskip
 
@@
 require some improvements to the tutorial pages
-(Section~\ref{sec:outlook}).
+(Section~\ref{sec:outlook}).\medskip
 
@@
   features in an unlabelled set of images will be allowed to count new
-  samples.
+  samples. The pilot data of the crowd-sourcing experiment show that,
+  despite the large scatter of the raw track counts obtained by the
+  citizen scientists, their pairwise ratio estimates scatter evenly
+  around the expert opinion. This suggests that the collective wisdom
+  of the entire group of students is greater than that of the
+  individual analysts. When applied to a group that has achieved a
+  good level of training and experience, crowd sourcing should improve
+  the precision and accuracy of fission track data.
 \item Length measurements. The geochron@home archive already

### 2025-07-21 geochron@home/R/main.R

@@ -26,14 +26,18 @@ PAsPA <- NULL
 total_counts_in_A0 <- c(N00=0,nP=0,nA=0)
 total_counts <- c(N1=0,N2=0)
 total_area <- c(A0=0,A1=0,A2=0)
+px2um2 <- mean(results[,'area_pixels']/results[,'area_mm2'])/1e6
+PAsPA <- matrix(NA,nrow=length(grains),ncol=2)
+colnames(PAsPA) <- c('PA','sPA')
+rhoP <- rhoA <- rep(NA,length(grains))
 for (i in seq_along(grains)){
     grain <- grains[i]
     ROIP <- PieterResults[[grain]]$ROI
     ROIA <- AndyResults[[grain]]$ROI
     ROI0 <- polygon_intersection(ROIP,ROIA)
-    A1 <- polygon_area(ROIP)
-    A2 <- polygon_area(ROIA)
-    A0 <- polygon_area(ROI0)
+    A1 <- polygon_area(ROIP,px2um2=px2um2)
+    A2 <- polygon_area(ROIA,px2um2=px2um2)
+    A0 <- polygon_area(ROI0,px2um2=px2um2)
     total_area <- total_area + c(A0,A1,A2)
     xyP <- PieterResults[[grain]]$counts
     xyA <- AndyResults[[grain]]$counts
@@ -45,11 +49,11 @@ for (i in seq_along(grains)){
     counts_in_A0 <- getN00(xyPinA0,xyAinA0)
     N00 <- counts_in_A0['N00']
     total_counts_in_A0 <- total_counts_in_A0 + counts_in_A0
-    PA <- (N1*A2)/(N2*A1)
-    sPA <- PA * sqrt(1/N1+1/N2-2*N00/(N1*N2))
-    PAsPA <- rbind(PAsPA,c(PA,sPA))
+    rhoP[i] <- 100*N1/A1 # x 1e-6 cm-2
+    rhoA[i] <- 100*N2/A2 # x 1e-6 cm-2
+    PAsPA[i,'PA'] <- rhoP[i]/rhoA[i]
+    PAsPA[i,'sPA'] <- PAsPA[i,'PA'] * sqrt(1/N1+1/N2-2*N00/(N1*N2))
 }
-colnames(PAsPA) <- c('PA','sPA')
 
 #### 2. Plot all the superimposed ROIs and counts for PV and AC ####
[…]

### 2025-07-21 geochron@home/R/polygon_overlap.R

@@ -1,7 +1,7 @@
 # This set of functions was written by ChatGPT!
 
 # Function to compute the area of a polygon using the Shoelace formula
-polygon_area <- function(vertices) {
+polygon_area <- function(vertices,px2um2=1) {
   n <- nrow(vertices)
   area <- 0
   for (i in 1:(n-1)) {
@@ -12,7 +12,7 @@ polygon_area <- function(vertices) {
   area <- area +
       (vertices[n,1] * vertices[1,2]) -
       (vertices[1,1] * vertices[n,2])
-  return(abs(area) / 2)
+  return(abs(area) / 2 / px2um2)
 }
 
 # Function to check if a point is inside

### 2025-07-21 geochron@home/R/helper.R
Further implementation of Rex suggestions

@@ -239,6 +239,8 @@ compare_grains <- function(results,grain1,grain2,plot=TRUE,...){
              ylab=paste0('grain ',grain2),
              pch=21,bg=bg,...)
     }
+    rho <- cor(out$x)[1,2]
+    mtext(line=-1,text=paste0('correlation=',format(round(rho,2),nsmall=2)))
     out
 }
 
@@ -253,4 +255,19 @@ radialcrowd <- function(results,grain1,grain2,from=NA,to=NA,t0=NA,...){
     graphics::mtext(paste0('PV ratio = ',signif(adminratio,3)),line=-2,cex=0.8)
     zs_admin <- provenance:::x2zs(tail(radialdat$x,1),from=from,to=to,t0=t0)
     IsoplotR:::plot_radial_points(zs_admin,pch=21,bg='blue')
+}
+
+add_table <- function(grouped_list,sigdig=1,nsmall=1){
+    x1 <- 73
+    x2 <- 78
+    x3 <- 87
+    y <- length(grouped_list)+1
+    text(x=x1,y=y,labels='n',xpd=NA,pos=4)
+    text(x=x2,y=y,labels='mean',xpd=NA,pos=4)
+    text(x=x3,y=y,labels='s.d',xpd=NA,pos=4)
+    for (i in seq_along(grouped_list)){
+        text(x=x1,y=i,labels=length(grouped_list[[i]]),xpd=NA,pos=4)
+        text(x=x2,y=i,labels=format(round(mean(grouped_list[[i]]),sigdig),nsmall=nsmall),xpd=NA,pos=4)
+        text(x=x3,y=i,labels=format(round(sd(grouped_list[[i]]),sigdig),nsmall=nsmall),xpd=NA,pos=4)
+    }
 }

### 2025-07-21 geochron@home/R/main.R

@@ -142,8 +142,10 @@ dev.off()
 grain1 <- 23
 grain2 <- 25
 trustworthy_results <- clean_results(results)
-pdf(file='../output/radialcrowd.pdf',width=12,height=4)
-p1 <- par(mfrow=c(1,3),mar=c(3,3,1,1),mgp=c(2,1,0))
+pdf(file='../output/radialcrowd.pdf',width=8,height=5)
+layout(rbind(c(1,1,2),
+             c(1,1,3)))
+p1 <- par(mar=c(3,3,1,10),mgp=c(2,1,0))
 # box plot
 grouped_list <- split(trustworthy_results$count, trustworthy_results$index)
 boxplot(grouped_list,horizontal=TRUE,
@@ -152,8 +154,10 @@ boxplot(grouped_list,horizontal=TRUE,
 add_admin_count_to_boxplot(trustworthy_results,grouped_list)
 axis(1)
 legend('topleft',legend='a)',bty='n',cex=1.5,adj=c(1,0))
+# table
+add_table(grouped_list)
 # scatter plot
-p2 <- par(cex=0.8,mar=c(3,2.5,1,1),xpd=NA)
+p2 <- par(cex=0.8,mar=c(3,2.5,1,1),xpd=NA,bty='n')
 compare_grains(trustworthy_results,grain1,grain2)
 legend('topleft',legend='b)',bty='n',cex=1.2,adj=c(1,0))
 # radial plot

### 2025-07-21 geochron@home/manuscript/paper.tex
Finished Figure 3

@@
 %\begin{figure}[!ht]
-{ \centering \includegraphics[width=\linewidth]{radialcrowd.pdf}
-  \captionof{figure}{a) box plots of 25 crowd-sourced sets of fission
-    track counts for 69 analysts; b) comparison of grains~23 and 25,
-    which were chosen because of their high counts and variability; c)
-    radial plot of the pairwise ratios of the fission track counts for
-    grains~23 and 25. The results of analyst PV are shown in
-    blue.\medskip}
+{ \centering \includegraphics[width=15cm]{radialcrowd.pdf}
+  \captionof{figure}{a) Box plots and summary statistics of numbers of
+    spontaneous fission tracks counted in each of 25 grains by 68
+    novice counters (geology students) all using the same ROIs. Each
+    student counted tracks in a subset of the grains. b) The track
+    counts for Grain 23 plotted against those for Grain 25 for 37
+    students who counted both of those grains. c) Radial plot of the
+    ratios of the counts in panel b) using the precisions given by
+    Equation~\ref{eq:N12}. In each panel, the counts made by PV are
+    added as blue dots.\medskip}
   \label{fig:radialcrowd}

### 2025-07-21 geochron@home/R/helper.R

@@ -238,9 +238,9 @@ compare_grains <- function(results,grain1,grain2,plot=TRUE,...){
              xlab=paste0('grain ',grain1),
              ylab=paste0('grain ',grain2),
              pch=21,bg=bg,...)
+        rho <- format(round(cor(out$x)[1,2],2),nsmall=2)
+        mtext(line=-1,text=paste0('correlation=',rho),cex=0.7)
     }
-    rho <- cor(out$x)[1,2]
-    mtext(line=-1,text=paste0('correlation=',format(round(rho,2),nsmall=2)))
     out
 }
 
@@ -251,23 +251,20 @@ radialcrowd <- function(results,grain1,grain2,from=NA,to=NA,t0=NA,...){
     admincount1 <- results$count[results$user_id==1 & results$index==grain1]
     admincount2 <- results$count[results$user_id==1 & results$index==grain2]
     adminratio <- admincount1/admincount2
-    graphics::mtext(paste0('pooled ratio = ',signif(pooledratio,3)),line=-1,cex=0.8)
-    graphics::mtext(paste0('PV ratio = ',signif(adminratio,3)),line=-2,cex=0.8)
+    graphics::mtext(paste0('pooled ratio = ',signif(pooledratio,3)),line=-1,cex=0.7)
+    graphics::mtext(paste0('PV ratio = ',signif(adminratio,3)),line=-2,cex=0.7)
     zs_admin <- provenance:::x2zs(tail(radialdat$x,1),from=from,to=to,t0=t0)
     IsoplotR:::plot_radial_points(zs_admin,pch=21,bg='blue')
 }
 
-add_table <- function(grouped_list,sigdig=1,nsmall=1){
-    x1 <- 73
-    x2 <- 78
-    x3 <- 87
+add_table <- function(grouped_list,sigdig=1,nsmall=1,x=c(70,75,85)){
     y <- length(grouped_list)+1
-    text(x=x1,y=y,labels='n',xpd=NA,pos=4)
-    text(x=x2,y=y,labels='mean',xpd=NA,pos=4)
-    text(x=x3,y=y,labels='s.d',xpd=NA,pos=4)
+    text(x=x[1],y=y,labels='n',xpd=NA,pos=4)
+    text(x=x[2],y=y,labels='mean',xpd=NA,pos=4)
+    text(x=x[3],y=y,labels='s.d',xpd=NA,pos=4)
[…]

### 2025-07-21 geochron@home/R/main.R

@@ -142,10 +142,10 @@ dev.off()
 grain1 <- 23
 grain2 <- 25
 trustworthy_results <- clean_results(results)
-pdf(file='../output/radialcrowd.pdf',width=8,height=5)
-layout(rbind(c(1,1,2),
-             c(1,1,3)))
-p1 <- par(mar=c(3,3,1,10),mgp=c(2,1,0))
+pdf(file='../output/radialcrowd.pdf',width=7,height=5)
+layout(rbind(c(1,1,1,2,2),
+             c(1,1,1,3,3)))
+p1 <- par(mar=c(3,3,1,8),mgp=c(2,1,0))
 # box plot
 grouped_list <- split(trustworthy_results$count, trustworthy_results$index)
 boxplot(grouped_list,horizontal=TRUE,
@@ -157,12 +157,12 @@ legend('topleft',legend='a)',bty='n',cex=1.5,adj=c(1,0))
 # table
 add_table(grouped_list)
 # scatter plot
-p2 <- par(cex=0.8,mar=c(3,2.5,1,1),xpd=NA,bty='n')
+p2 <- par(mar=c(3,2.5,1,1),xpd=NA,bty='n')
 compare_grains(trustworthy_results,grain1,grain2)
-legend('topleft',legend='b)',bty='n',cex=1.2,adj=c(1,0))
+legend('topleft',legend='b)',bty='n',cex=1.0,adj=c(1,0))
 # radial plot
-radialcrowd(trustworthy_results,grain1,grain2,from=0.4,to=2.5,t0=1)
-legend('topleft',legend='c)',bty='n',cex=1.2,adj=c(1,0))
+radialcrowd(trustworthy_results,grain1,grain2,from=0.4,to=3.0,t0=1)
+legend('topleft',legend='c)',bty='n',cex=1.0,adj=c(1,0))
 par(p2)
 par(p1)
 dev.off()

### 2025-07-21 geochron@home/manuscript/paper.tex
Finished geochron@home supplement

@@
 \begin{abstract}
-Fission track geochronology is based on the visual analysis of optical
-images. This visual process is prone to observer bias. Fission track
-datasets are currently reported as small data tables. The
+Fission track thermochronology is based on the visual analysis of
+optical images. This visual process is prone to observer bias. Fission
+track datasets are currently reported as small data tables. The
 interpretation of these tables requires a high degree of trust between
@@
 reviewed fission track counts can be used to train and improve machine
-learning algorithms for automated fission track analysis. We hope that
-other geochronological methods will follow the fission track
-community's lead in FAIR data processing. This would benefit all the
-other Earth Science disciplines that depend on geochronological data.
+learning algorithms for automated fission track analysis. We invite
+other geochronological methods to follow the fission track community's
+lead in FAIR data processing. This would benefit all the other Earth
+Science disciplines that depend on geochronological data.
 \end{abstract}
@@
 \item Post-processing. The fission track data can either be downloaded
-  as a flat data table of counts and area, or as a \texttt{.json} file
-  containing the locations of all the counted tracks. geochron@home
-  does not provide any tools to post-process these files. They are
-  meant to be passed on to other tools such as spreadsheet
-  applications or IsoplotR \citep{vermeesch2018c}.
+  as a flat data table of counts and areas, or as a \texttt{.json}
+  file containing the locations of all the counted
+  tracks. geochron@home does not provide any tools to post-process
+  these files. They are meant to be passed on to other tools such as
+  spreadsheet applications or IsoplotR \citep{vermeesch2018c}.
 
@@
 fission track data. The next sections of this paper will illustrate
-these four different applications with real world examples.\medskip
+these applications with real world examples.\medskip
 
@@
 sample. Figure~\ref{fig:AvP} illustrates this with two sets of fission
-track density estimates (in tracks per \textmu{m}\textsuperscript{2})
-for the same sample of Mount Dromedary apatite \citep{green1985b},
-analysed by two users (PV and AC).\medskip
+track density estimates for the same sample of Mount Dromedary apatite
+\citep{green1985b}, analysed by two users (PV and AC).\medskip
 
-Let $N_1$ and $N_2$ be the numbers of spontaneous tracks counted by PV
-and AC over areas $A_1$ and $A_2$, respectively. Then their estimated
-track densities are give by $\hat{\rho}_1 = N_1/A_1$ and $\hat{\rho}_2
-= N_2/A_2$. In this situation when the two areas overlap, $N_1$ and
-$N_2$ cannot be treated as independent Poisson counts because the two
-analysts will count some of the same tracks. It can be shown that,
-under simple (ideal) assumptions, the uncertainty of the ratio of the
-estimated track densities is given approximately by
+Let $N_1$ and $N_2$ be the numbers of spontaneous tracks counted by
+the two users over areas $A_1$ and $A_2$, respectively. Then their
+estimated track densities are given by $\hat{\rho}_1 = N_1/A_1$ and
+$\hat{\rho}_2 = N_2/A_2$. In this situation when the two areas
+overlap, $N_1$ and $N_2$ cannot be treated as independent Poisson
+counts because the two analysts will count some of the same tracks. It
+can be shown that, under simple (ideal) assumptions, the uncertainty
+of the ratio of the estimated track densities is given approximately
+by
 \begin{equation}
@@
 
-The radial plot of Figure~\ref{fig:AvP}b contains two sets of counts,
-with $\sum N_{1} = 686$ and $\sum N_{2} = 679$, and a weighted mean
-track density ratio $\hat{\rho}_{1}/\hat{\rho}_{1} = 0.94$ with
-relative standard error 0.013. PV counted 600 tracks in those common
-areas $A_0$, of which 549 were also counted by AC. Conversely, AC
-counted 622 tracks of which 549 were also counted by PV. The ratio of
-the two analysts' track density estimates based just on the common
-area is therefore 600/622 = 0.965 with relative standard error 0.016,
-calculated from Equation~\ref{eq:NO}. This is close to the weighted
-mean ratio of 0.94 (Figure~\ref{fig:AvP}b), and is slightly nearer to
-1. It indicates that PV under-counts the Mount Dromedary apatite by
-3.5\% relative to AC.  The existence of `observer bias' is well
-documented \citep{tamer2025}.  It is one of the reasons why fission
-track analysis is often done relative to age standards: observer bias
-does not have to be a problem provided that it is consistent between
-grains, and between samples.\medskip
+The combined data plots of Figures~\ref{fig:AvP}b and c contain two
+sets of counts, with $\sum N_{1} = 686$ and $\sum N_{2} = 679$, and a
+weighted mean track density ratio $\hat{\rho}_{1}/\hat{\rho}_{1} =
+0.94$ with relative standard error 0.013. PV counted 600 tracks in
+those common areas $A_0$, of which 549 were also counted by
+AC. Conversely, AC counted 622 tracks of which 549 were also counted
+by PV. The ratio of the two analysts' track density estimates based
+just on the common area is therefore 600/622 = 0.965 with relative
+standard error 0.016, calculated from Equation~\ref{eq:NO}. This is
+close to the weighted mean ratio of 0.94 (Figure~\ref{fig:AvP}c), and
+is slightly nearer to 1. It indicates that PV under-counts the Mount
+Dromedary apatite by 3.5\% relative to AC.  The existence of `observer
+bias' is well documented \citep{tamer2025}.  It is one of the reasons
+why fission track analysis is often done relative to age standards:
+observer bias does not have to be a problem provided that it is
+consistent between grains, and between samples.\medskip
 
@@
 provides a quick and easy mechanism to train novice users in the art
-of fission track analysis. The tutorial pages are in their infancy,
+of fission track analysis. The tutorial pages are in their infancy
 and offer only a limited degree of interactivity. Users can click on
@@
     student counted tracks in a subset of the grains. b) The track
-    counts for Grain 23 plotted against those for Grain 25 for 37
+    counts for Grain 25 plotted against those for Grain 23 for 37
     students who counted both of those grains. c) Radial plot of the
@@
 
-A 1984 interlaboratory comparison study \citep{miller1985} showed that
+An interlaboratory comparison study by \citet{miller1985} showed that
 the average of several fission track age estimates is closer to the
 known age of mineral standards than the age obtained by any individual
-observer. Unfortunately, routine measurement of fission track samples
-by multiple analysts is prohibitively expensive in a normal laboratory
-environment. geochron@home changes this, by bringing fission track
+observer. Routine measurement of fission track samples by multiple
+analysts is prohibitively expensive in a normal laboratory
+environment. geochron@home changes this by bringing fission track
 analysis to a proverbial `county fair' of citizen scientists.\medskip
@@
 be expected because of the differing numbers of tracks due to varying
-areas and uranium contents of the grains. But the variation between
-counts within grains (shown by the standard deviations) is due
-entirely to differences between the students' recognising and counting
-exactly the same tracks. These standard deviations increase with the
-mean number of tracks and are considerable in size, being on average
-about 40\% of the mean.  Also, the vast majority of students counted
-fewer tracks than PV did, often many fewer. In 21 of the 25 grains,
-PV's count is above the upper quartile of the students', and on 2
-occasions someone counted no tracks at all.  With expert trained
-counters one would expect much smaller differences between counts.
-Nevertheless there are many repeated counts also.\medskip
+areas and U contents of the grains. But the variation between counts
+within grains (shown by the standard deviations) is due entirely to
+differences between the students' recognising and counting exactly the
+same tracks. These standard deviations increase with the mean number
+of tracks and are considerable in size, being on average about 40\% of
+the mean.  Also, the vast majority of students counted fewer tracks
+than PV did, often many fewer. In 21 of the 25 grains, PV's count is
+above the upper quartile of the students', and on two occasions
+someone counted no tracks at all.  With expert trained counters one
+would expect much smaller differences between counts.  Nevertheless
+there are many repeated counts also.\medskip
 
 Figure~\ref{fig:radialcrowd}b shows a scatter plot of pairs of counts
-for two grains made by 33 students who counted both of them. They
-average around 23 and 30 tracks per grain compared with PV's pair of
-44 and 63, which is highlighted in blue. There is a positive
+for two grains made by 37 students who counted both of them. They
+average around 29 and 30 tracks per grain compared with PV's pair of
+52 and 50, which is highlighted in blue. There is a positive
 correlation between the pairs, consistent with systematic observer
@@
 with Poisson variation.  Figure~\ref{fig:radialcrowd}c shows a radial
-plot of the ratios of the counts in Panel b, using the precisions
+plot of the ratios of the counts in Panel~b, using the precisions
 given by
@@
 Figure~\ref{fig:AvP}b, where we were comparing the ratios of counts
-made by 2 analysts on each of 25 grains. Here we are comparing the
-ratios of counts for the same 2 grains made by each of 33
+made by two analysts on each of 25 grains. Here we are comparing the
+ratios of counts for the same two grains made by each of 37
 analysts. There is no theoretical justification for
@@
 expert counters one would expect less variation than that. However, in
-spite of the under-counting by the students, the pooled the students'
-ratio of 0.76 is quite close to PV's ratio of 0.70. This happens quite
+spite of the under-counting by the students, the pooled students'
+ratio of 0.76 is close to PV's ratio of 0.70. This happens quite
 often in this data set, but not always, and there are some pairs of
@@
 
-geochron@home has been in development for 10 years and
-remains a work in progress. Planned improvements include:
+geochron@home has been in development for a decade and remains a work
+in progress. Planned improvements include:
 
@@
   features in an unlabelled set of images will be allowed to count new
-  samples. The pilot data of the crowd-sourcing experiment show that,
-  despite the large scatter of the raw track counts obtained by the
-  citizen scientists, their pairwise ratio estimates scatter evenly
-  around the expert opinion. This suggests that the collective wisdom
-  of the entire group of students is greater than that of the
-  individual analysts. When applied to a group that has achieved a
-  good level of training and experience, crowd sourcing should improve
-  the precision and accuracy of fission track data.
+  samples. Section~\ref{sec:crowdsourcing} shows that, despite the
+  large scatter of the raw track counts obtained by the citizen
+  scientists, their pairwise ratio estimates scatter evenly around the
+  expert opinion. This suggests that the collective wisdom of the
+  entire group of students is greater than that of the individual
+  analysts. When applied to a group that has achieved a good level of
+  training and experience, crowd sourcing should improve the precision
+  and accuracy of fission track data.
 \item Length measurements. The geochron@home archive already
@@
   the horizontal etch rates along the c-axis and parallel to it
-  \citep{donelick1993}. Functions will be added measure and visualise
-  these type of data in geochron@home.
+  \citep{donelick1993}. Functions will be added to measure and
+  visualise this type of data in geochron@home.
 \item Machine learning. Data science is experiencing an artificial
@@
 
-This paper introduced geochron@home, a software platform for
-FAIR fission track analysis. We demonstrated four different
-applications of this platform using real data. Putting the FAIR data
-paradigm into practice, all the imagery, counts and source code for
-this paper are publicly available. Using these resources, the reader
-can reproduce all the results that were presented in this publication.
-We hope that other geochronologists will follow this example.  FAIR
-data promises to address the reproducibility crisis in science
+This paper introduced geochron@home, a software platform for FAIR
+fission track analysis. We demonstrated four different applications of
+this platform using real data. Putting the FAIR data paradigm into
+practice, all the imagery, counts and source code for this paper are
+publicly available. Using these resources, the reader can reproduce
+all the results that were presented in this publication.  We encourage
+other geochronologists to follow this example.  FAIR data promises to
+address the reproducibility crisis in science
 \citep{miyakawa2020}.\medskip
@@
 in fission track images analysis will be used to update the count data
-for published datasets, improved their accuracy.\medskip
+for published datasets, improving their accuracy.\medskip
 
@@
 
-With the development of geochron@home, fission track geochronology is
-once again ahead of the pack in terms of FAIR data analysis.
+With the development of geochron@home, fission track thermochronology
+is once again ahead of the pack in terms of FAIR data analysis.
 geochron@home currently only stores images and counts. This is enough
@@
   from \url{https://github.com/pvermees/geochron-at-home} (last
-  access: April~2, 2025).  The raw data (imagery) are available at the
+  access: July~21, 2025).  The raw data (imagery) are available at the
   geochron@home archive (\url{https://github.com/pvermees/GaHa}, last
-  access: April~3, 2025).  R-scripts to reproduce the figures are
+  access: July~21, 2025).  R-scripts to reproduce the figures are
   provided in the supplementary information
   (\url{https://github.com/pvermees/supplements}, last access:
-  April~3, 2025).}
+  July~21, 2025).}
 
@@
   provided the samples, prepared the training data and counted fission
-  tracks. Rex Galbraith derived Equation~\ref{eq:NO}.}
+  tracks. Rex Galbraith derived Equation~\ref{eq:NO} and verified all
+  the other calculations.}
 

### 2025-07-21 geochron@home/R/main.R

@@ -71,15 +71,15 @@ op <- par(mar=c(4,4,.5,.5),mfrow=c(1,3),mgp=c(2.5,1,0),cex=1.0)
 plotROIs(PieterResults[['4648']],AndyResults[['4648']])
 legend('topleft','a)',bty='n')
 plot(rhoA,rhoP,type='n',
-     xlab=expression(hat(rho)[AC]*"(cm"^2%*%"10"^-6*")"),
-     ylab=expression(hat(rho)[PV]*"(cm"^2%*%"10"^-6*")"),
+     xlab=expression(hat(rho)[AC]*"(cm"^-2%*%"10"^-6*")"),
+     ylab=expression(hat(rho)[PV]*"(cm"^-2%*%"10"^-6*")"),
      bty='n',log='xy',xlim=c(0.1,2),ylim=c(0.1,2))
 abline(a=0,b=1)
 points(rhoA,rhoP,type='p')
 legend('topleft','b)',bty='n')
 mtext(text='log ratios PV/AC',line=-1)
 mtext(text=paste0('mean = ',signif(mean(log(PAsPA[,'PA'])),3)),line=-2)
-mtext(text=paste0('st.dev = ',signif(sd(log(PAsPA[,'PA'])),3)),line=-3)
+mtext(text=paste0('s.d = ',signif(sd(log(PAsPA[,'PA'])),3)),line=-3)
 mtext(text=paste0('n = ',nrow(PAsPA)),line=-4)
 par(mar=c(4,4,3,1))
 PAradial(PAsPA,cex=1.0,spacing=1.2)

### 2025-07-23 geochron@home/manuscript/paper.tex
Updated author contributions

@@
 \Author[2]{Jiangping}{He}
+\Author[1]{Rex}{Galbraith}
 \Author[3]{Andrew}{Carter}
-\Author[1]{Rex}{Galbraith}
 
@@
 
-\authorcontribution{Pieter Vermeesch designed the study, acquired the
-  funding, counted fission tracks and wrote the paper. Jiangping He
-  created geochron@home. Tim Band expanded geochron@home and wrote the
-  accompanying microscope image acquisition software. Andrew Carter
-  provided the samples, prepared the training data and counted fission
-  tracks. Rex Galbraith derived Equation~\ref{eq:NO} and verified all
-  the other calculations.}
+\authorcontribution{PV designed the study, acquired the funding and
+  counted fission tracks. JH created geochron@home. TB expanded
+  geochron@home and wrote the accompanying microscope image
+  acquisition software. RG derived Equation~\ref{eq:NO} and verified
+  the other calculations. AC provided the samples, prepared the
+  training data and counted fission tracks. PV and RG wrote the paper
+  with input from the other authors.}
 

### 2025-07-26 geochron@home/manuscript/paper.tex
Added summary tables to geochron@home paper

@@
 fission tracks; and (4) the `wisdom of crowds' in fission track
-identification.\medskip
+ratio estimation.\medskip
 
@@
 other geochronological methods to follow the fission track community's
-lead in FAIR data processing. This would benefit all the other Earth
-Science disciplines that depend on geochronological data.
+lead in FAIR data processing. This would benefit all the Earth Science
+disciplines that depend on geochronological data.
 \end{abstract}
@@
 just on the common area is therefore 600/622 = 0.965 with relative
-standard error 0.016, calculated from Equation~\ref{eq:NO}. This is
+standard error 0.018, calculated from Equation~\ref{eq:NO}. This is
 close to the weighted mean ratio of 0.94 (Figure~\ref{fig:AvP}c), and
@@
 
+We here present the results of a type of `crowd-sourcing' experiment
+that was carried out at UCL as part of an undergraduate course in
+isotope geology. 70 students were asked to create an account on the
+geochron@home server by selecting a unique username and
+password. After spending a few minutes to complete the compulsory
+tutorial (Section~\ref{sec:tutorial}), the students were asked to
+count fission tracks in randomly assigned image stacks of Mount
+Dromedary apatite. Each student was required to analyse at least 15
+grains, using pre-defined ROIs. In a matter of hours, the students
+amassed a dataset of 37,331 fission track counts in 34 grains.  This
+large dataset was parsed into separate data files --- one for each
+student --- which they processed during an assessed practical
+exercise.\medskip
+
+The raw counts for 68 students (two of the students were unable to
+complete the exercise) are given in Table~\ref{tab:crowdtable} of the
+Appendix, for the same selection of 25 grains that were analysed by AC
+and PV in Section~\ref{sec:private}, along with corresponding counts
+made by PV. Everyone counted tracks in the same ROIs.
 Figure~\ref{fig:radialcrowd}a shows box plots and summary statistics
-for the raw counts made by 68 students (two of the students were
-unable to complete the assignment) for the same selection of 25 grains
-that were analysed by AC and PV in Section~\ref{sec:private}. Everyone
-counted tracks in the same ROIs, and counts made by PV for those ROIs
-have been added as blue dots.  Variation between the mean values is to
-be expected because of the differing numbers of tracks due to varying
-areas and U contents of the grains. But the variation between counts
-within grains (shown by the standard deviations) is due entirely to
-differences between the students' recognising and counting exactly the
-same tracks. These standard deviations increase with the mean number
-of tracks and are considerable in size, being on average about 40\% of
-the mean.  Also, the vast majority of students counted fewer tracks
-than PV did, often many fewer. In 21 of the 25 grains, PV's count is
-above the upper quartile of the students', and on two occasions
-someone counted no tracks at all.  With expert trained counters one
-would expect much smaller differences between counts.  Nevertheless
-there are many repeated counts also.\medskip
+of these counts for each grain.  Variation between the mean values is
+to be expected because of the differing numbers of tracks due to
+varying areas and U contents of the grains. But the variation between
+counts within grains (shown by the standard deviations) is due
+entirely to differences between the students’ recognising and counting
+exactly the same tracks. These standard deviations increase with the
+mean number of tracks and are considerable in size, being on average
+about 40\% of the mean. With expert trained counters one would expect
+much smaller differences between counts.  Furthermore, the vast
+majority of students counted fewer tracks than PV did, often many
+fewer, and on two occasions someone counted no tracks at all. PV's
+count is always above the students' median and nearly always above the
+upper quartile.  Table~\ref{tab:crowdtable} (looking along rows) shows
+just 7 students with average counts within 10\% of PV's, and 19 with
+average counts less than half of his.\medskip
 
 Figure~\ref{fig:radialcrowd}b shows a scatter plot of pairs of counts
-for two grains made by 37 students who counted both of them. They
+for two grains made by 36 students who counted both of them. They
 average around 29 and 30 tracks per grain compared with PV's pair of
@@
 similarly low in the other, and vice versa).  Comparing other pairs of
-grains shows similar results, with correlations varying between 0.3
+grains shows similar results, with correlations varying between 0.4
 and 0.9. In all cases we have looked at, the residual variation, after
@@
 spite of the under-counting by the students, the pooled students'
-ratio of 0.76 is close to PV's ratio of 0.70. This happens quite
-often in this data set, but not always, and there are some pairs of
-grains where the students' pooled ratio differs somewhat from
-PV's.\medskip
+ratio of 0.97 is close to PV's ratio of 1.04. This happens quite often
+in this data set, but not always, and there are some pairs of grains
+where the students' pooled ratio differs somewhat from PV's.  The
+logcontrast between the student ratios and PV's ratios for all 300
+pairs of grains follows a normal distribution with a mean of 0.0 and
+standard deviation 0.23 (Table~\ref{tab:logcontrasts}). This indicates
+that, on the whole, the student estimates of the (log)ratios are
+unbiassed with respect to the expert opinion of PV.
+\medskip
 
@@
 probabilities, reflecting the ambiguity of some fission track
-datasets.\medskip
-
-Figure~\ref{fig:crowdsourcing} also shows that some students counted
-the tails of the fission tracks rather than their etch pits, despite
-being told the opposite in the tutorial. Fixing this issue will
-require some improvements to the tutorial pages
+datasets. Figure~\ref{fig:crowdsourcing} also shows that some students
+counted the tails of the fission tracks rather than their etch pits,
+despite being told the opposite in the tutorial. Fixing this issue
+will require some improvements to the tutorial pages
 (Section~\ref{sec:outlook}).\medskip
@@
 
-%\appendix
-%\section{}
+\appendix
+\section{Summary tables for the crowd-sourcing experiment}
 
-%\appendix{}
+\begingroup
+\captionof{table}{ Summary table of the crowd sourcing results for the
+  different users (rows) and grains (columns).  The last row is the
+  ratio of the median count to PV's count (user~1) for each grain.
+  The last column is the ratio of the weighted mean count over grains
+  compared with PV for the same grains.
+}
+\label{tab:crowdtable}
+\renewcommand{\arraystretch}{0.7}
+\begin{scriptsize}
+  \input{../output/crowdtable.txt}
+\end{scriptsize}
+\endgroup
+
+\clearpage
+
+\begingroup
+\captionof{table}{
+  Logcontrasts between the pooled count ratio of the
+  students and PV for each pair of 25 grains in
+  Table~\ref{tab:crowdtable}.
+}
+\label{tab:logcontrasts}\medskip
+\renewcommand{\arraystretch}{0.7}
+\begin{scriptsize}
+  \input{../output/logcontrasts.txt}
+\end{scriptsize}
+\endgroup
 

### 2025-07-26 geochron@home/R/helper.R

@@ -76,6 +76,7 @@ PAradial <- function(PAsPA,cex=1.0,spacing=1.0){
     if (fit$p.value<0.05){
         mtext(text=disptit,line=-spacing,cex=cex)
     }
+    invisible(fit)
 }
 
 # gets ROI from rois.json
@@ -246,6 +247,8 @@ compare_grains <- function(results,grain1,grain2,plot=TRUE,...){
 
 radialcrowd <- function(results,grain1,grain2,from=NA,to=NA,t0=NA,...){
     radialdat <- compare_grains(results,grain1,grain2,plot=FALSE)
+    duplicates <- duplicated(radialdat)
+    radialdat$x[duplicates,] <- jitter(radialdat$x[duplicates,])
     provenance:::radialplot.counts(radialdat,title=FALSE,from=from,to=to,t0=t0,...)
     pooledratio <- sum(radialdat$x[,1])/sum(radialdat$x[,2])
     admincount1 <- results$count[results$user_id==1 & results$index==grain1]
@@ -255,6 +258,7 @@ radialcrowd <- function(results,grain1,grain2,from=NA,to=NA,t0=NA,...){
     graphics::mtext(paste0('PV ratio = ',signif(adminratio,3)),line=-2,cex=0.7)
     zs_admin <- provenance:::x2zs(tail(radialdat$x,1),from=from,to=to,t0=t0)
     IsoplotR:::plot_radial_points(zs_admin,pch=21,bg='blue')
+    invisible(radialdat)
 }
 
 add_table <- function(grouped_list,sigdig=1,nsmall=1,x=c(70,75,85)){
@@ -267,4 +271,86 @@ add_table <- function(grouped_list,sigdig=1,nsmall=1,x=c(70,75,85)){
         text(x=x[2],y=i,labels=format(round(mean(grouped_list[[i]]),sigdig),nsmall=nsmall),xpd=NA,pos=4)
         text(x=x[3],y=i,labels=format(round(sd(grouped_list[[i]]),sigdig),nsmall=nsmall),xpd=NA,pos=4)
     }
+}
+
+crowdtable <- function(trustworthy_results){
+    count_table <- reshape(
+        trustworthy_results[,c('user_id','index','count')],
+        direction = "wide",
+        idvar = "user_id",
+        timevar = "index",
+        sep = ""
+    )
+    area_table <- reshape(
+        trustworthy_results[,c('user_id','index','area_pixels')],
+        direction = "wide",
+        idvar = "user_id",
+        timevar = "index",
[…]

### 2025-07-26 geochron@home/R/main.R

@@ -161,7 +161,10 @@ p2 <- par(mar=c(3,2.5,1,1),xpd=NA,bty='n')
 compare_grains(trustworthy_results,grain1,grain2)
 legend('topleft',legend='b)',bty='n',cex=1.0,adj=c(1,0))
 # radial plot
-radialcrowd(trustworthy_results,grain1,grain2,from=0.4,to=3.0,t0=1)
+radialgrain12 <- radialcrowd(trustworthy_results,grain1,grain2,from=0.4,to=3.0,t0=1)
+ng <- nrow(radialgrain12$x)
+pooledratio12crowd <- sum(radialgrain12$x[-ng,1])/sum(radialgrain12$x[-ng,2])
+pooledratio12PV <- radialgrain12$x[ng,1]/radialgrain12$x[ng,2]
 legend('topleft',legend='c)',bty='n',cex=1.0,adj=c(1,0))
 par(p2)
 par(p1)
@@ -170,3 +173,30 @@ dev.off()
 # generalised linear fit to crowd-sourced data
 fit <- glm(count ~ index + user_id, data=trustworthy_results, family="poisson")
 mswd <- summary(fit)$deviance/summary(fit)$df.residual
+
+# summary tables
+lst <- crowdtable(trustworthy_results)
+counts2latex(lst,destination='../output/crowdtable.txt')
+
+ordered_counts <- lst$tab
+indices <- colnames(ordered_counts)
+ni <- length(indices)
+crowdratios <- matrix(0,nrow=ni,ncol=ni,dimnames=list(indices,indices))
+PVratios <- matrix(0,nrow=ni,ncol=ni,dimnames=list(indices,indices))
+for (i1 in indices){
+    for (i2 in indices){
+        paired <- which(!(is.na(ordered_counts[,i1]) | is.na(ordered_counts[,i2])))[-1]
+        crowdratios[i1,i2] <- sum(ordered_counts[paired,i1])/sum(ordered_counts[paired,i2])
+        PVratios[i1,i2] <- ordered_counts[1,i1]/ordered_counts[1,i2]
+    }
+}
+logcontrast <- log(crowdratios)-log(PVratios)
+formatted <- as.dist(apply(logcontrast, MARGIN = c(1, 2), FUN = function(x) sprintf("%.2f", x)))
+
+pdf(file='../output/logcontrasts.pdf',width=4.5,height=4.5)
+op <- par(mar=c(5,5,0,0))
[…]

### 2025-07-30 geochron@home/manuscript/paper.tex
Removed logcontrasts

@@
 in this data set, but not always, and there are some pairs of grains
-where the students' pooled ratio differs somewhat from PV's.  The
-logcontrast between the student ratios and PV's ratios for all 300
-pairs of grains follows a normal distribution with a mean of 0.0 and
-standard deviation 0.23 (Table~\ref{tab:logcontrasts}). This indicates
-that, on the whole, the student estimates of the (log)ratios are
-unbiassed with respect to the expert opinion of PV.
+where the students' pooled ratio differs somewhat from PV's.
 \medskip
@@
 
-\clearpage
-
-\begingroup
-\captionof{table}{
-  Logcontrasts between the pooled count ratio of the
-  students and PV for each pair of 25 grains in
-  Table~\ref{tab:crowdtable}.
-}
-\label{tab:logcontrasts}\medskip
-\renewcommand{\arraystretch}{0.7}
-\begin{scriptsize}
-  \input{../output/logcontrasts.txt}
-\end{scriptsize}
-\endgroup
-
 %\noappendix

### 2025-07-30 geochron@home/R/helper.R

@@ -207,7 +207,10 @@ clean_results <- function(results,ACgrains){
     good_grains <- results$index %in% AC_ids
     good_counts <- (timely_submission & is_trustworthy & uses_default_roi) | is_admin
     keep <- good_grains & good_counts
-    results[keep,]
+    out <- results[keep,]
+    out$index <- as.factor(out$index)
+    out$user_id <- as.factor(out$user_id)
+    out
 }
 
 add_admin_count_to_boxplot <- function(trustworthy_results,grouped_list){
@@ -331,26 +334,4 @@ counts2latex <- function(lst,destination){
     }
     cat('\\end{tabular}\n',file=file_conn)
     close(file_conn)    
-}
-
-dist2latex <- function(d,destination){
-    file_conn <- file(destination, open = "w")
-    cat('\\begin{tabular}{',file=file_conn)
-    n <- attr(d,"Size")
-    cat(paste0(rep('r@{~}',n),collapse=''),file=file_conn)
-    cat('}\n',file=file_conn)
-    dmat <- as.matrix(d)
-    labels <- attr(d,"Labels")
-    cat(paste0(c('',labels[-n]),collapse=' & '), file = file_conn)
-    cat(' \\cr\n',file=file_conn)
-    for (i in 2:n){
-        cat(paste0(labels[i],' & ',collapse=''),file = file_conn)
-        values <- sprintf('%.2f',dmat[i,1:(i-1)])
-        shortvals <- gsub('0\\.', '.', values)
-        cat(paste0(shortvals,collapse=' & '),file = file_conn)
-        cat(paste0(rep('',n-i),collapse=' & '),file = file_conn)
-        cat(' \\cr\n',file = file_conn)
-    }
-    cat('\\end{tabular}\n',file=file_conn)
-    close(file_conn)
 }

### 2025-07-30 geochron@home/R/main.R

@@ -171,32 +171,9 @@ par(p1)
 dev.off()
 
 # generalised linear fit to crowd-sourced data
-fit <- glm(count ~ index + user_id, data=trustworthy_results, family="poisson")
+fit <- glm(count ~ user_id + index, data=trustworthy_results, family="poisson")
 mswd <- summary(fit)$deviance/summary(fit)$df.residual
 
 # summary tables
 lst <- crowdtable(trustworthy_results)
 counts2latex(lst,destination='../output/crowdtable.txt')
-
-ordered_counts <- lst$tab
-indices <- colnames(ordered_counts)
-ni <- length(indices)
-crowdratios <- matrix(0,nrow=ni,ncol=ni,dimnames=list(indices,indices))
-PVratios <- matrix(0,nrow=ni,ncol=ni,dimnames=list(indices,indices))
-for (i1 in indices){
-    for (i2 in indices){
-        paired <- which(!(is.na(ordered_counts[,i1]) | is.na(ordered_counts[,i2])))[-1]
-        crowdratios[i1,i2] <- sum(ordered_counts[paired,i1])/sum(ordered_counts[paired,i2])
-        PVratios[i1,i2] <- ordered_counts[1,i1]/ordered_counts[1,i2]
-    }
-}
-logcontrast <- log(crowdratios)-log(PVratios)
-formatted <- as.dist(apply(logcontrast, MARGIN = c(1, 2), FUN = function(x) sprintf("%.2f", x)))
-
-pdf(file='../output/logcontrasts.pdf',width=4.5,height=4.5)
-op <- par(mar=c(5,5,0,0))
-hist(logcontrast,main='',col='white')
-par(op)
-dev.off()
-
-dist2latex(formatted,destination='../output/logcontrasts.txt')

### 2025-08-20 geochron@home/manuscript/paper.tex
Trimmed down crowdsourcing section

@@
     Andrew Carter). a) the track counts and ROIs of both users for
-    grain 4648, shown in blue (PV) and red (AC); b) comparison of the
+    grain~3, shown in blue (PV) and red (AC); b) comparison of the
     track densities for PV and AC for all 25 grains analysed by the
@@
 
-%\begin{figure}[!ht]
-{ \centering \includegraphics[width=15cm]{radialcrowd.pdf}
-  \captionof{figure}{a) Box plots and summary statistics of numbers of
-    spontaneous fission tracks counted in each of 25 grains by 68
-    novice counters (geology students) all using the same ROIs. Each
-    student counted tracks in a subset of the grains. b) The track
-    counts for Grain 25 plotted against those for Grain 23 for 37
-    students who counted both of those grains. c) Radial plot of the
-    ratios of the counts in panel b) using the precisions given by
-    Equation~\ref{eq:N12}. In each panel, the counts made by PV are
-    added as blue dots.\medskip}
-  \label{fig:radialcrowd}
-}%\end{figure}
+\begingroup
+\captionof{table}{ Summary table of the crowd sourcing results for the
+  different users (rows) and grains (columns).  The last row is the
+  ratio of the median count to PV's count (user~1) for each grain.
+  The last column is the ratio of the weighted mean count over grains
+  compared with PV for the same grains.
+}
+\label{tab:crowdtable}
+\renewcommand{\arraystretch}{0.7}
+\begin{scriptsize}
+  \input{../output/crowdtable.txt}
+\end{scriptsize}
+\endgroup
 
@@
 \citep{galton1907a}. Similar effects are seen in fission track
-geochronology.\medskip
-
-An interlaboratory comparison study by \citet{miller1985} showed that
-the average of several fission track age estimates is closer to the
-known age of mineral standards than the age obtained by any individual
-observer. Routine measurement of fission track samples by multiple
-analysts is prohibitively expensive in a normal laboratory
-environment. geochron@home changes this by bringing fission track
-analysis to a proverbial `county fair' of citizen scientists.\medskip
+geochronology. An interlaboratory comparison study by
+\citet{miller1985} showed that the average of several fission track
+age estimates is closer to the known age of mineral standards than the
+age obtained by any individual observer. Inspired by these previous
+experiments, we used geochron@home to bring fission track analysis to
+a proverbial `county fair' of citizen scientists.\medskip
 
@@
 large dataset was parsed into separate data files --- one for each
-student --- which they processed during an assessed practical
+student --- which they processed during an assessed programming
 exercise.\medskip
@@
 The raw counts for 68 students (two of the students were unable to
-complete the exercise) are given in Table~\ref{tab:crowdtable} of the
-Appendix, for the same selection of 25 grains that were analysed by AC
-and PV in Section~\ref{sec:private}, along with corresponding counts
-made by PV. Everyone counted tracks in the same ROIs.
-Figure~\ref{fig:radialcrowd}a shows box plots and summary statistics
-of these counts for each grain.  Variation between the mean values is
-to be expected because of the differing numbers of tracks due to
-varying areas and U contents of the grains. But the variation between
-counts within grains (shown by the standard deviations) is due
-entirely to differences between the students’ recognising and counting
-exactly the same tracks. These standard deviations increase with the
-mean number of tracks and are considerable in size, being on average
-about 40\% of the mean. With expert trained counters one would expect
-much smaller differences between counts.  Furthermore, the vast
-majority of students counted fewer tracks than PV did, often many
-fewer, and on two occasions someone counted no tracks at all. PV's
-count is always above the students' median and nearly always above the
-upper quartile.  Table~\ref{tab:crowdtable} (looking along rows) shows
-just 7 students with average counts within 10\% of PV's, and 19 with
-average counts less than half of his.\medskip
+complete the exercise) are given in Table~\ref{tab:crowdtable}, for
+the same selection of 25 grains that were analysed by AC and PV in
+Section~\ref{sec:private}, along with corresponding counts made by
+PV. Everyone counted tracks in the same ROIs. The vast majority of
+students counted fewer tracks than PV did, often many fewer, and on
+two occasions someone counted no tracks at all. PV's count is always
+above the students' median and nearly always above the upper quartile.
+Table~\ref{tab:crowdtable} (looking along rows) shows just 7 students
+with average counts within 10\% of PV's, and 18 with average counts
+less than half of his.\medskip
 
-Figure~\ref{fig:radialcrowd}b shows a scatter plot of pairs of counts
-for two grains made by 36 students who counted both of them. They
-average around 29 and 30 tracks per grain compared with PV's pair of
-52 and 50, which is highlighted in blue. There is a positive
-correlation between the pairs, consistent with systematic observer
-effects (i.e., people who count a low value in one grain tend to count
-similarly low in the other, and vice versa).  Comparing other pairs of
-grains shows similar results, with correlations varying between 0.4
-and 0.9. In all cases we have looked at, the residual variation, after
-allowing for systematic differences between students, is consistent
-with Poisson variation.  Figure~\ref{fig:radialcrowd}c shows a radial
-plot of the ratios of the counts in Panel~b, using the precisions
-given by
-\begin{equation}
-  \frac{\mbox{se}(N_1/N_2)}{N_1/N_2}
-  \approx
-  \sqrt{
-    \frac{1}{N_1} + \frac{1}{N_2}
-  }
-  \label{eq:N12}
-\end{equation}
-
-The context here is rather different from that in
-Figure~\ref{fig:AvP}b, where we were comparing the ratios of counts
-made by two analysts on each of 25 grains. Here we are comparing the
-ratios of counts for the same two grains made by each of 37
-analysts. There is no theoretical justification for
-Equation~\ref{eq:N12} here, but it still provides a useful benchmark
-for assessing variation between the ratios. In
-Figure~\ref{fig:radialcrowd}c, that variation is consistent with
-Poisson variation (as it is for most other pairs of grains) but with
-expert counters one would expect less variation than that. However, in
-spite of the under-counting by the students, the pooled students'
-ratio of 0.97 is close to PV's ratio of 1.04. This happens quite often
-in this data set, but not always, and there are some pairs of grains
-where the students' pooled ratio differs somewhat from PV's.
-\medskip
+However, in spite of the under-counting by the students, the
+\emph{ratios} of the track counts between different grains are
+generally quite close to PV's ratios, especially when comparing grains
+with similar track densities. This is apparent from
+Table~\ref{tab:crowdtable} (looking along columns).
+Figure~\ref{fig:23vs25}b shows a scatter plot of pairs of counts for
+two grains (23 and 25) made by 36 students who counted both of
+them. They average around 29 and 30 tracks per grain compared with
+PV's pair of 52 and 50, which is highlighted in blue. There is a
+positive correlation between the pairs, consistent with systematic
+observer effects (i.e., people who count a low value in one grain tend
+to count similarly low in the other, and vice versa).  Comparing other
+pairs of grains shows similar results, with correlations varying
+between 0.4 and 0.9. In all cases we have looked at, the residual
+variation, after allowing for systematic differences between students,
+is consistent with Poisson variation.
 
@@
 downloaded as a \texttt{.json} file and inspected in detail.
-Figure~\ref{fig:crowdsourcing} is a two-dimensional histogram for the
-x-y positions of all the track positions generated by the students in
-the two apatite crystals. Visual comparison of the histogram with the
+Figure~\ref{fig:23vs25}b) shows two-dimensional histograms for the x-y
+positions of all the track positions generated by the students in the
+two apatite crystals. Visual comparison of the histogram with the
 optical image confirms that the students unanimously identified the
@@
 probabilities, reflecting the ambiguity of some fission track
-datasets. Figure~\ref{fig:crowdsourcing} also shows that some students
+datasets. Figure~\ref{fig:23vs25} also shows that some students
 counted the tails of the fission tracks rather than their etch pits,
@@
 %\begin{figure}[!ht]
-{  \centering
-  \includegraphics[width=9cm]{4649vs4673.pdf}
-  \captionof{figure}{
-    a,b) Optical images in transmitted light of 
-    Mount Dromedary apatites 4649 and 4673 in the crowd-sourcing experiment.
-    c,d) Two-dimensional histogram of the track locations for the
-    two grains, as identified by the citizen scientists.
+{ \centering \includegraphics[width=9cm]{23vs25.pdf}
+  \captionof{figure}{ a) The track counts for grain~25 plotted against
+    those for grain~23 for 37 students who counted both of those
+    grains. b) Optical images in transmitted light of Mount Dromedary
+    apatites 23 and 25 in the crowd-sourcing experiment.  c,d)
+    Two-dimensional histogram of the track locations for the two
+    grains, as identified by the citizen scientists.
     \medskip}
-  \label{fig:crowdsourcing}
+  \label{fig:23vs25}
 }%\end{figure}
@@
 
-\appendix
-\section{Summary tables for the crowd-sourcing experiment}
-
-\begingroup
-\captionof{table}{ Summary table of the crowd sourcing results for the
-  different users (rows) and grains (columns).  The last row is the
-  ratio of the median count to PV's count (user~1) for each grain.
-  The last column is the ratio of the weighted mean count over grains
-  compared with PV for the same grains.
-}
-\label{tab:crowdtable}
-\renewcommand{\arraystretch}{0.7}
-\begin{scriptsize}
-  \input{../output/crowdtable.txt}
-\end{scriptsize}
-\endgroup
-
 %\noappendix

### 2025-08-20 geochron@home/R/helper.R

@@ -244,15 +244,8 @@ compare_grains <- function(results,grain1,grain2,plot=TRUE,...){
              xlab=paste0('grain ',grain1),
              ylab=paste0('grain ',grain2),
              pch=21,bg=bg,...)
-        h1 <- poisson.test(out$x[iadmin,1])
-        h2 <- poisson.test(out$x[iadmin,2])
-        arrows(x0=c(out$x[iadmin,1],h1$conf.int[1]),
-               x1=c(out$x[iadmin,1],h1$conf.int[2]),
-               y0=c(h2$conf.int[1],out$x[iadmin,2]),
-               y1=c(h2$conf.int[2],out$x[iadmin,2]),
-               angle=90,length=0.05,code=3)
         rho <- format(round(cor(out$x)[1,2],2),nsmall=2)
-        mtext(line=-1,text=paste0('correlation=',rho),cex=0.7)
+        mtext(line=-1,text=paste0('correlation=',rho),cex=1.0)
     }
     out
 }

### 2025-08-20 geochron@home/R/main.R

@@ -70,7 +70,7 @@ for (grain in grains){
 par(op)
 dev.off()
 
-#### 3. Compare PV and AC's results for grain 4648 ####
+#### 3. Compare PV and AC's results for grain 3 (a.k.a. 4648) ####
 pdf(file='../output/AvP.pdf',width=15,height=5)
 op <- par(mar=c(4,4,.5,.5),mfrow=c(1,3),mgp=c(2.5,1,0),cex=1.0)
 plotROIs(PieterResults[['4648']],AndyResults[['4648']])
@@ -109,29 +109,27 @@ for (grain in grains){
 dev.off()
 
 #### 5. Show crowdsourcing results for two selected grains ####
-grain1 <- '4649'
-grain2 <- '4673'
-pdf(file="../output/4649vs4673.pdf",onefile=FALSE)
+grain1 <- list(number='4685',index=23) # list(number='4682',index=39)
+grain2 <- list(number='4674',index=25) # list(number='4680',index=43)
+pdf(file="../output/23vs25.pdf",onefile=FALSE)
 op <- par(mar=c(4,4,0,1))
 image1 <- plotimage(idir='../screenshots',
-                    grain=allgrains[[grain1]]$grain) +
-    labs(tag = "a)") +
+                    grain=allgrains[[grain1$number]]$grain) +
+    labs(tag = grain1$index) +
     theme(legend.position="right",
           legend.direction="vertical",
           legend.key.height=unit(1.0, "cm"))
-counts1 <- plotcounts(allgrains[[grain1]],roiss) +
-    labs(tag = "c)") +
+counts1 <- plotcounts(allgrains[[grain1$number]],roiss) +
     theme(legend.position="bottom",
           legend.direction="horizontal",
           legend.key.width=unit(1.0, "cm"))
 image2 <- plotimage(idir='../screenshots',
-                    grain=allgrains[[grain2]]$grain) +
-    labs(tag = "b)") +
+                    grain=allgrains[[grain2$number]]$grain) +
+    labs(tag = grain2$index) +
     theme(legend.position="right",
           legend.direction="vertical",
           legend.key.height=unit(1.0, "cm"))
-counts2 <- plotcounts(allgrains[[grain2]],roiss) +
-    labs(tag = "d)") +
[…]

### 2025-08-21 geochron@home/manuscript/paper.tex
Changes to outlook section

@@
 \begingroup
-\captionof{table}{ Summary table of the crowd sourcing results for the
+\captionof{table}{Summary table of the crowd sourcing results for the
   different users (rows) and grains (columns).  The last row is the
-  ratio of the median count to PV's count (user~1) for each grain.
-  The last column is the ratio of the weighted mean count over grains
-  compared with PV for the same grains.
-}
+  ratio of the median count to PV's count for each grain.  The last
+  column is the ratio of the pooled count over grains compared with PV
+  for the same grains.}
 \label{tab:crowdtable}
@@
 age estimates is closer to the known age of mineral standards than the
-age obtained by any individual observer. Inspired by these previous
+age obtained by individual observers. Inspired by these previous
 experiments, we used geochron@home to bring fission track analysis to
@@
 password. After spending a few minutes to complete the compulsory
-tutorial (Section~\ref{sec:tutorial}), the students were asked to
-count fission tracks in randomly assigned image stacks of Mount
-Dromedary apatite. Each student was required to analyse at least 15
-grains, using pre-defined ROIs. In a matter of hours, the students
-amassed a dataset of 37,331 fission track counts in 34 grains.  This
-large dataset was parsed into separate data files --- one for each
-student --- which they processed during an assessed programming
+tutorial (Section~\ref{sec:tutorial}), they were asked to count
+fission tracks in randomly assigned image stacks of Mount Dromedary
+apatite. Each student was required to analyse at least 15 grains,
+using pre-defined ROIs. In a matter of hours, the students amassed a
+dataset of 37,331 fission track counts in 34 grains.  This large
+dataset was parsed into separate data files --- one for each student
+--- which they processed during an assessed programming
 exercise.\medskip
@@
 with similar track densities.  This is apparent from
-Table~\ref{tab:crowdtable} (looking along columns).
-Figure~\ref{fig:23vs25}b shows a scatter plot of pairs of counts for
-two grains (23 and 25) made by 36 students who counted both of
-them. They average around 29 and 30 tracks per grain compared with
-PV's pair of 52 and 50, which is highlighted in blue. There is a
-positive correlation between the pairs, consistent with systematic
-observer effects (i.e., people who count a low value in one grain tend
-to count similarly low in the other, and vice versa).  Comparing other
-pairs of grains shows similar results, with correlations varying
-between 0.4 and 0.9. In all cases we have looked at, the residual
-variation, after allowing for systematic differences between students,
-is consistent with Poisson variation.
+Table~\ref{tab:crowdtable} (looking along columns).  In grains
+containing few tracks (left hand side of the table), the median of the
+student counts is within 20\% of PV's counts.  However, this gradually
+levels off to 50\% at higher track densities (right hand side of the
+table). Figure~\ref{fig:23vs25}b shows a scatter plot of pairs of
+counts for two grains (23 and 25) made by 36 students who counted both
+of them. The two grains were chosen because they have similar track
+counts for both PV and the median student. They both have a median of
+28 tracks per grain compared with PV's pair of 52 and 50, which is
+highlighted in blue. There is a positive correlation between the
+pairs, consistent with systematic observer effects (i.e., people who
+count a low value in one grain tend to count similarly low in the
+other, and vice versa).  Comparing other pairs of grains shows similar
+results, with correlations varying between 0.4 and 0.9. Pairs of
+grains with similar track densities are more strongly correlated than
+those with dissimilar track densities. In all cases we have looked at,
+the residual variation, after allowing for systematic differences
+between students, is consistent with Poisson variation.\medskip
 
@@
 Figure~\ref{fig:23vs25}b) shows two-dimensional histograms for the x-y
-positions of all the track positions generated by the students in the
-two apatite crystals. Visual comparison of the histogram with the
-optical image confirms that the students unanimously identified the
-most obvious semitracks, which contain a clearly visible etch pit and
+positions of all the track positions generated by the students in
+grains~23 and 25. Visual comparison of the histogram with the optical
+image confirms that the students unanimously identified the most
+obvious semitracks, which contain a clearly visible etch pit and
 tail. Shorter and fainter tracks received fewer clicks. Datasets like
@@
 %\begin{figure}[!ht]
-{ \centering \includegraphics[width=9cm]{23vs25.pdf}
+{ \centering \includegraphics[width=15cm]{23vs25.pdf}
   \captionof{figure}{ a) The track counts for grain~25 plotted against
@@
     grains. b) Optical images in transmitted light of Mount Dromedary
-    apatites 23 and 25 in the crowd-sourcing experiment.  c,d)
-    Two-dimensional histogram of the track locations for the two
+    apatites 23 and 25 in the crowd-sourcing experiment, along with
+    two-dimensional histogram of the track locations for the two
     grains, as identified by the citizen scientists.
@@
 \begin{enumerate}
-\item Interactive tutorial pages. To ensure that novice users do not
-  count the tails but the etch pits of fission tracks, we will add a
-  `quiz' to the tutorial pages. Only users who click on the `correct'
-  features in an unlabelled set of images will be allowed to count new
-  samples. Section~\ref{sec:crowdsourcing} shows that, despite the
-  large scatter of the raw track counts obtained by the citizen
-  scientists, their pairwise ratio estimates scatter evenly around the
-  expert opinion. This suggests that the collective wisdom of the
-  entire group of students is greater than that of the individual
-  analysts. When applied to a group that has achieved a good level of
-  training and experience, crowd sourcing should improve the precision
-  and accuracy of fission track data.
 \item Length measurements. The geochron@home archive already
@@
   visualise this type of data in geochron@home.
+\item Interactive tutorial pages and embedded quality
+  control. Section~\ref{sec:crowdsourcing} shows that the
+  crowd-sourcing idea is not yet ready for general use. Two
+  improvements will be made to make its results more reliable. First,
+  a `quiz' to the tutorial pages to ensure that novice users do not
+  count the tails but the etch pits of fission tracks. Only users who
+  click on the `correct' features in an unlabelled set of images will
+  be allowed to count new samples. Second, future citizen scientists
+  will occasionally be served reference images of known track
+  density. Their counts for these `test' images will be used to
+  identify trustworthy analysts (top of Table~\ref{tab:crowdtable}),
+  and flag unreliable ones (bottom of Table~\ref{tab:crowdtable}).
 \item Machine learning. Data science is experiencing an artificial
@@
   counts. This data format is close to the form in which data are
-  treated within an AI algorithm.\medskip
-
-  Once trained on historical data, AI algorithms can be used to count
-  fission tracks automatically. Following the model of Fission Track
-  Studio \citep{gleadow2009, gleadow2019}, machine learning can be
-  used to reverse the fission track counting process. Instead of
-  asking users to count the fission tracks in a sample, the software
-  can ask them to check the results proposed by an AI algorithm, and
-  to remove any features that are \emph{not} fission tracks.\medskip
-
-  Regardless of whether fission tracks were counted manually or with a
-  machine, the value of the geochron@home archive remains the
-  same. It is important to document data so that samples can be
-  reanalysed in the future, for example when a new and improved
+  treated within an AI algorithm.
+\item Inverse counting. Once trained on historical data, AI algorithms
+  can be used to count fission tracks automatically. Following the
+  model of Fission Track Studio \citep{gleadow2009, gleadow2019},
+  machine learning can be used to reverse the fission track counting
+  process. Instead of asking users to count the fission tracks in a
+  sample, the software can ask them to check the results proposed by
+  an AI algorithm, and to remove any features that are \emph{not}
+  fission tracks.  Regardless of whether fission tracks were counted
+  manually or with a machine, the value of the geochron@home archive
+  remains the same. It is important to document data so that samples
+  can be reanalysed in the future, for example when a new and improved
   generation of machine learning algorithms becomes available.
@@
 
-These four improvements will be made by ourselves pending additional
+These improvements will be made by ourselves pending additional
 funding. However, because geochron@home is free and open software, we
@@
 
-Another advantage of the geochron@home wokflow is the separation of
+Another advantage of the geochron@home workflow is the separation of
 image acquisition and image analysis. This separation reduces the

### 2025-08-21 geochron@home/manuscript/paper.tex
Side caption for Table 1 in g@h paper

@@
 inspection by other users; (3) to create tutorials for new students of
-the fission track method; and (4) to generate crowd-sourced fission
-track data, by serving randomly selected selections of images to
-citizen scientists. We illustrate these four applications with
-examples that demonstrate (1) geochron@home's ability to compare and
-combine fission track counts for multiple users within a lab group;
-(2) the value of the geochron@home archive in the peer review system;
-(3) the use of simple tutorials in teaching novice users how to count
-fission tracks; and (4) the `wisdom of crowds' in fission track
-ratio estimation.\medskip
+the fission track method; and (4) to serve randomly selected
+selections of images to citizen scientists. We illustrate these four
+applications with examples that demonstrate (1) geochron@home's
+ability to compare and combine fission track counts for multiple users
+within a lab group; (2) the value of the geochron@home archive in the
+peer review system; (3) the use of simple tutorials in teaching novice
+users how to count fission tracks; and (4) the opportunities and
+challenges of crowd-sourced fission track analysis.\medskip
 
@@
 digital microscopy has moved fission track data acquisition from the
-objective lens of a microscope to the computer screen
+ocular lenses of a microscope to the computer screen
 \citep{gleadow2009, vanranst2019, gleadow2019}. Ongoing developments
@@
   directory. Because Fission Track Studio stores its results in an
-  \texttt{.xml} format, a second conversion script is needed to
+  \texttt{.xml} format, a second conversion script was created to
   translate those results into an equivalent \texttt{.json} format.
@@
 
-\begingroup
-\captionof{table}{Summary table of the crowd sourcing results for the
-  different users (rows) and grains (columns).  The last row is the
-  ratio of the median count to PV's count for each grain.  The last
-  column is the ratio of the pooled count over grains compared with PV
-  for the same grains.}
-\label{tab:crowdtable}
-\renewcommand{\arraystretch}{0.7}
-\begin{scriptsize}
-  \input{../output/crowdtable.txt}
-\end{scriptsize}
-\endgroup
-
 In 1906, Sir Francis Galton visited a county fair in which a contest
@@
 However, in spite of the under-counting by the students, the
-\emph{ratios} of the track counts between different grains are
-generally quite close to PV's ratios, especially when comparing grains
-with similar track densities.  This is apparent from
-Table~\ref{tab:crowdtable} (looking along columns).  In grains
-containing few tracks (left hand side of the table), the median of the
-student counts is within 20\% of PV's counts.  However, this gradually
-levels off to 50\% at higher track densities (right hand side of the
-table). Figure~\ref{fig:23vs25}b shows a scatter plot of pairs of
+\emph{ratios} of the track counts between different grains are often
+close to PV's ratios, especially when comparing grains with similar
+track densities.  This is apparent from Table~\ref{tab:crowdtable}
+(looking along columns).  In grains containing few tracks (left hand
+side of the table), the median of the student counts is within 20\% of
+PV's counts.  However, this gradually levels off to 50\% at higher
+track densities (right hand side of the
+table). Figure~\ref{fig:23vs25}a shows a scatter plot of pairs of
 counts for two grains (23 and 25) made by 36 students who counted both
@@
 other, and vice versa).  Comparing other pairs of grains shows similar
-results, with correlations varying between 0.4 and 0.9. Pairs of
-grains with similar track densities are more strongly correlated than
-those with dissimilar track densities. In all cases we have looked at,
-the residual variation, after allowing for systematic differences
-between students, is consistent with Poisson variation.\medskip
+results, with correlations varying between 0.4 and 0.9. In all cases
+we have looked at, the residual variation, after allowing for
+systematic differences between students, is consistent with Poisson
+variation.\medskip
 
@@
 downloaded as a \texttt{.json} file and inspected in detail.
-Figure~\ref{fig:23vs25}b) shows two-dimensional histograms for the x-y
+Figure~\ref{fig:23vs25}b shows two-dimensional histograms for the x-y
 positions of all the track positions generated by the students in
-grains~23 and 25. Visual comparison of the histogram with the optical
-image confirms that the students unanimously identified the most
+grains~23 and 25. Visual comparison of the histograms with the optical
+images confirms that the students unanimously identified the most
 obvious semitracks, which contain a clearly visible etch pit and
@@
 (Section~\ref{sec:outlook}).\medskip
+
+\noindent\begin{minipage}{.65\linewidth}
+\renewcommand{\arraystretch}{0.7}
+\begin{scriptsize}
+  \input{../output/crowdtable.txt}
+\end{scriptsize}
+\end{minipage}
+\begin{minipage}{.35\linewidth}
+\captionof{table}{Summary table of the crowd sourcing results for the
+  different users (rows) and grains (columns).  The last row is the
+  ratio of the median count to PV's count for each grain.  The last
+  column is the ratio of the pooled count over grains compared with PV
+  for the same grains.}
+\label{tab:crowdtable}
+\end{minipage}
 
@@
   improvements will be made to make its results more reliable. First,
-  a `quiz' to the tutorial pages to ensure that novice users do not
-  count the tails but the etch pits of fission tracks. Only users who
-  click on the `correct' features in an unlabelled set of images will
-  be allowed to count new samples. Second, future citizen scientists
-  will occasionally be served reference images of known track
-  density. Their counts for these `test' images will be used to
-  identify trustworthy analysts (top of Table~\ref{tab:crowdtable}),
-  and flag unreliable ones (bottom of Table~\ref{tab:crowdtable}).
+  a `quiz' will be added to the tutorial pages to ensure that novice
+  users do not count the tails but the etch pits of fission
+  tracks. Only users who click on the `correct' features in an
+  unlabelled set of images will be allowed to count new
+  samples. Second, future citizen scientists will occasionally be
+  served reference images of known track density. Their counts for
+  these `test' images will be used to identify trustworthy analysts
+  (top of Table~\ref{tab:crowdtable}) and flag unreliable ones (bottom
+  of Table~\ref{tab:crowdtable}).
 \item Machine learning. Data science is experiencing an artificial
-  intelligence revolution that has already started to transform the
-  fission track method \citep{nachtergaele2020}.  Convolutional neural
-  networks must be trained with example data. geochron@home
-  is ideally suited for this task. Section~\ref{sec:crowdsourcing}
-  showed how the collective wisdom of multiple fission track analysts
-  can label fission track images with probabilities rather than
+  intelligence (AI) revolution that has already started to transform
+  the fission track method \citep{nachtergaele2020}.  Convolutional
+  neural networks must be trained with example data. geochron@home is
+  ideally suited for this task. Section~\ref{sec:crowdsourcing} showed
+  how the opinion of multiple fission track analysts can be combined
+  to label fission track images with probabilities rather than
   counts. This data format is close to the form in which data are
@@
 practice, all the imagery, counts and source code for this paper are
-publicly available. Using these resources, the reader can reproduce
-all the results that were presented in this publication.  We encourage
-other geochronologists to follow this example.  FAIR data promises to
-address the reproducibility crisis in science
-\citep{miyakawa2020}.\medskip
+publicly available on the geochron@home archive. Using these
+resources, the reader can reproduce all the results that were
+presented in this publication.  We encourage other geochronologists to
+follow this example.  FAIR workflows promise to address the
+reproducibility crisis in science \citep{miyakawa2020}.\medskip
 
@@
   from \url{https://github.com/pvermees/geochron-at-home} (last
-  access: July~21, 2025).  The raw data (imagery) are available at the
-  geochron@home archive (\url{https://github.com/pvermees/GaHa}, last
-  access: July~21, 2025).  R-scripts to reproduce the figures are
-  provided in the supplementary information
+  access: August~21, 2025).  The raw data (imagery) are available at
+  the geochron@home archive (\url{https://github.com/pvermees/GaHa},
+  last access: August~21, 2025).  R-scripts to reproduce the figures
+  are provided in the supplementary information
   (\url{https://github.com/pvermees/supplements}, last access:
-  July~21, 2025).}
+  August~21, 2025).}
 
@@
   geochron@home and wrote the accompanying microscope image
-  acquisition software. RG derived Equation~\ref{eq:NO} and verified
-  the other calculations. AC provided the samples, prepared the
-  training data and counted fission tracks. PV and RG wrote the paper
-  with input from the other authors.}
+  acquisition software. RG derived Equation~\ref{eq:NO}, designed
+  Table~\ref{tab:crowdtable} and verified the other calculations. AC
+  provided the samples, prepared the training data and counted fission
+  tracks. PV and RG wrote the paper with input from the other
+  authors.}
 

### 2025-09-27 geochron@home/manuscript/paper.tex
Created smaller version of the table

@@
 just on the common area is therefore 600/622 = 0.965 with relative
-standard error 0.018, calculated from Equation~\ref{eq:NO}. This is
-close to the weighted mean ratio of 0.94 (Figure~\ref{fig:AvP}c), and
-is slightly nearer to 1. It indicates that PV under-counts the Mount
-Dromedary apatite by 3.5\% relative to AC.  The existence of `observer
-bias' is well documented \citep{tamer2025}.  It is one of the reasons
-why fission track analysis is often done relative to age standards:
-observer bias does not have to be a problem provided that it is
-consistent between grains, and between samples.\medskip
+standard error 0.018, calculated from Equation~\ref{eq:NO}, with $N_1
+= 686$, $N_2 = 679$ and $N_0 = 549$.  This is close to the weighted
+mean ratio of 0.94 (Figure~\ref{fig:AvP}c), and is slightly nearer to
+1. It indicates that PV under-counts the Mount Dromedary apatite by
+3.5\% relative to AC.  The existence of `observer bias' is well
+documented \citep{tamer2025}.  It is one of the reasons why fission
+track analysis is often done relative to age standards: observer bias
+does not have to be a problem provided that it is consistent between
+grains, and between samples.\medskip
 
@@
 
-\noindent\begin{minipage}{.65\linewidth}
-\renewcommand{\arraystretch}{0.7}
-\begin{scriptsize}
-  \input{../output/crowdtable.txt}
-\end{scriptsize}
-\end{minipage}
-\begin{minipage}{.35\linewidth}
-\captionof{table}{Summary table of the crowd sourcing results for the
-  different users (rows) and grains (columns).  The last row is the
-  ratio of the median count to PV's count for each grain.  The last
-  column is the ratio of the pooled count over grains compared with PV
-  for the same grains.}
+\begin{small}
+  \input{../output/shortcrowdtable.txt}\\
+\end{small}
+\captionof{table}{Counts of fission tracks in the same ROIs of 25
+  grains made by PV and 68 students. Only 20 grains and 20 students
+  are shown due to space constraints; the full table is provided in
+  the supplementary information.  Each student counted a subset of the
+  grains. The grains are listed in increasing order of PV's
+  counts. The last 3 columns give, for each student, the number of
+  grains counted $n_g$ and the weighted mean $r_p =
+  \sum\,w_i\,r_i/\sum\,w_i$ and standard deviation $s_p =
+  \sqrt{w_i\,(r_i-r_p)^2/(n_g − 1)}$ where $r_i$ is the ratio of the
+  student's count to PV's count for grain $i$ and $w_i$ is PV's count
+  for grain $i$. ($r_p$ is the same as the ratio of the student's
+  total count over the $n_g$ grains to PV's total count over the same
+  grains.)  The students are listed in decreasing order of $r_p$ . The
+  last 3 rows ($n$, $\bar{y}$ and $s_y$ ) give the sample size (number
+  of students), mean and standard deviation of the students' counts
+  for each grain.}
 \label{tab:crowdtable}
-\end{minipage}
+%\end{minipage}
 
@@
 { \centering \includegraphics[width=15cm]{23vs25.pdf}
-  \captionof{figure}{ a) The track counts for grain~25 plotted against
-    those for grain~23 for 37 students who counted both of those
-    grains. b) Optical images in transmitted light of Mount Dromedary
-    apatites 23 and 25 in the crowd-sourcing experiment, along with
-    two-dimensional histogram of the track locations for the two
-    grains, as identified by the citizen scientists.
+  \captionof{figure}{a) The track counts for grain~23 plotted against
+    those for grain 25 for 36 students who counted both of those
+    grains. The grey line through the origin has slope equal to the
+    ratio of the total counts for each grain. The blue dot shows PV's
+    counts for those two grains. b) Optical images in transmitted
+    light of Mount Dromedary apatites 23 and 25 in the crowd-sourcing
+    experiment, along with two-dimensional histogram of the track
+    locations for the two grains, as identified by the citizen
+    scientists.
     \medskip}

### 2025-09-27 geochron@home/R/helper.R

@@ -245,7 +245,17 @@ compare_grains <- function(results,grain1,grain2,plot=TRUE,...){
              ylab=paste0('grain ',grain2),
              pch=21,bg=bg,...)
         rho <- format(round(cor(out$x)[1,2],2),nsmall=2)
-        mtext(line=-1,text=paste0('correlation=',rho),cex=1.0)
+        mtext(line=0,text=paste0('correlation=',rho),cex=1.0)
+        PVratio <- out$x[iadmin,1]/out$x[iadmin,2]
+        mtext(line=-2,
+              text=paste0("PV's ratio=",
+                          format(round(PVratio,2),nsmall=2)),
+              cex=1.0)
+        pooled_ratio <- sum(results1$count[i1])/sum(results2$count[i2])
+        mtext(line=-3,
+              text=paste0("pooled ratio=",
+                          format(round(pooled_ratio,2),nsmall=2)),
+              cex=1.0)
     }
     out
 }
@@ -300,39 +310,66 @@ crowdtable <- function(trustworthy_results){
         res[order(as.integer(rownames(res))),
             order(as.integer(colnames(res)))]
     }
-    ordered_counts <- order_table(count_table,first=6) # 6 characters in "count."    
-    rowratio <- rep(NA,nrow(ordered_counts))
+    ordered_counts <- order_table(count_table,first=6) # 6 characters in "count."
+    ng <- rp <- sp <- rep(NA,nrow(ordered_counts))
     for (i in 1:nrow(ordered_counts)){
         good <- !is.na(ordered_counts[i,])
-        rowratio[i] <- sum(ordered_counts[i,good])/sum(ordered_counts[1,good])
+        ng[i] <- sum(good)
+        w <- ordered_counts[1,good]
+        r <- ordered_counts[i,good]/ordered_counts[1,good]
+        rp[i] <- sum(w*r)/sum(w)
+        sp[i] <- sqrt(sum(w*(r-rp[i])^2)/(ng[i]-1))
     }
-    colmedians <- apply(ordered_counts[-1,],2,'median',na.rm=TRUE)
-    colratio <- colmedians/ordered_counts[1,]
-    roworder <- c(1,1+order(rowratio[-1],decreasing=TRUE))
+    n <- colSums(!is.na(ordered_counts))
[…]

### 2025-09-27 geochron@home/R/main.R

@@ -146,7 +146,9 @@ trustworthy_results <- clean_results(results)
 #### 6. Scatter plot of crowdsourcing results ####
 pdf(file='../output/crowd-correlation.pdf',width=5,height=5)
 op <- par(mar=c(3,3,1,1),mgp=c(2,1,0),xpd=NA,bty='n')
-compare_grains(trustworthy_results,grain2$index,grain1$index)#,xlim=c(0,60),ylim=c(0,70))
+lim <- c(0,70)
+compare_grains(trustworthy_results,grain2$index,grain1$index,xlim=lim,ylim=lim)
+lines(x=lim,y=lim,col='grey50',lwd=1.5)
 par(op)
 dev.off()
 
@@ -172,3 +174,5 @@ mswd <- summary(fit)$deviance/summary(fit)$df.residual
 # summary tables
 lst <- crowdtable(trustworthy_results)
 counts2latex(lst,destination='../output/crowdtable.txt')
+counts2latex(lst,destination='../output/shortcrowdtable.txt',short=TRUE)
+

### 2025-09-28 geochron@home/manuscript/paper.tex
Added dot plot.

@@
 
+\begin{small}
+  \input{../output/shortcrowdtable.txt}\\
+\end{small}
+\captionof{table}{Counts of fission tracks in the same ROIs of 25
+  grains made by PV and 68 students. Only 20 students are shown due to
+  space constraints; the full table is provided in the supplementary
+  information.  Each student counted a subset of the grains. The
+  grains are listed in increasing order of PV's counts. The last 3
+  columns give, for each student, the number of grains counted $n_g$
+  and the weighted mean $r_p = \sum\,w_i\,r_i/\sum\,w_i$ and standard
+  deviation $s_p = \sqrt{\sum w_i\,(r_i-r_p)^2/(n_g − 1)}$
+  where $r_i$ is the ratio of the student's count to PV's count for
+  grain $i$ and $w_i$ is PV's count for grain $i$. ($r_p$ is the same
+  as the ratio of the student's total count over the $n_g$ grains to
+  PV's total count over the same grains.)  The students are listed in
+  decreasing order of $r_p$. $r_m$ is the ratio of the students'
+  median count per grain to PV's count.}
+\label{tab:crowdtable}
+
 The raw counts for 68 students (two of the students were unable to
-complete the exercise) are given in Table~\ref{tab:crowdtable}, for
-the same selection of 25 grains that were analysed by AC and PV in
-Section~\ref{sec:private}, along with corresponding counts made by
-PV. Everyone counted tracks in the same ROIs. The vast majority of
-students counted fewer tracks than PV did, often many fewer, and on
-two occasions someone counted no tracks at all. PV's count is always
-above the students' median and nearly always above the upper quartile.
+complete the exercise) are given in Table~\ref{tab:crowdtable} and
+Figure~\ref{fig:stripchart}, for the same selection of 25 grains that
+were analysed by AC and PV in Section~\ref{sec:private}, along with
+corresponding counts made by PV. Everyone counted tracks in the same
+ROIs. Variation between the mean values is to be expected because of
+the differing numbers of tracks due to varying areas and U contents of
+the grains. But the variation between counts within grains (shown by
+the standard deviations in Figure~\ref{fig:stripchart}) is due
+entirely to differences between the students' recognising and counting
+exactly the same tracks. These standard deviations increase with the
+mean number of tracks and are considerable in size, being on average
+about 40\% of the mean. With expert trained counters one would expect
+much smaller differences between counts.  Furthermore, the vast
+majority of students counted fewer tracks than PV did, often many
+fewer, and on two occasions someone counted no tracks at all. PV's
+count is always above the students' median and nearly always above the
+upper quartile (Figure~\ref{fig:stripchart}).
 Table~\ref{tab:crowdtable} (looking along rows) shows just 7 students
@@
 
-However, in spite of the under-counting by the students, the
-\emph{ratios} of the track counts between different grains are often
-close to PV's ratios, especially when comparing grains with similar
-track densities.  This is apparent from Table~\ref{tab:crowdtable}
-(looking along columns).  In grains containing few tracks (left hand
-side of the table), the median of the student counts is within 20\% of
-PV's counts.  However, this gradually levels off to 50\% at higher
-track densities (right hand side of the
-table). Figure~\ref{fig:23vs25}a shows a scatter plot of pairs of
-counts for two grains (23 and 25) made by 36 students who counted both
-of them. The two grains were chosen because they have similar track
-counts for both PV and the median student. They both have a median of
-28 tracks per grain compared with PV's pair of 52 and 50, which is
-highlighted in blue. There is a positive correlation between the
-pairs, consistent with systematic observer effects (i.e., people who
-count a low value in one grain tend to count similarly low in the
-other, and vice versa).  Comparing other pairs of grains shows similar
-results, with correlations varying between 0.4 and 0.9. In all cases
-we have looked at, the residual variation, after allowing for
-systematic differences between students, is consistent with Poisson
-variation.\medskip
+The rightmost three columns of Table~\ref{tab:crowdtable} contain
+summary statistics ($n_g$ , $r_p$ and $s_p$) giving, for each student,
+the number of grains counted and the weighted mean and standard
+deviation of the ratios of the student's count to PV's count for each
+grain (definitions provided in the caption). The true number of tracks
+in each ROI is unknown, so we are comparing the students with PV
+(remembering that PV's count is also subject to error, which is the
+same for each student). $r_p$ measures agreement with PV on average
+and $s_p$ measures consistency over grains. For good agreement $r_p$
+should be close to 1 and $s_p$ should be small.  Perfect agreement is
+$r_p = 1$ and $s_p = 0$. For comparison, an $r_p = 1.04$ and $s_p =
+0.69$ for PV and AC using just the counts in the common areas $A_0$
+from Section~\ref{sec:private}. With the caveat that the ROIs are
+different and PV's counts are different, this calculation shows that
+there is a group of 6 (possibly 9) students whose agreement with PV is
+comparable with AC's. This group does not include student 229 whose
+$r_p = 0.95$ but with a large $s_p = 1.62$.\medskip
+
+However, in spite of the under-counting by the students, the ratio of
+their pooled track counts ($r_p$) for two different grains is often
+close to PV's ratio, especially when comparing grains with similar
+track densities. This is illustrated in Figure~\ref{fig:23vs25}a for
+grains~23 and 25.  For any point, the ratio of the two counts is given
+by the slope of a line from the origin to that point. Although the
+individual slopes vary considerably, the line with slope equal to the
+ratio of the total counts for all students (which is 0.96) passes
+close to PV's pair of counts (52,50) shown by the blue dot (with ratio
+1.03). It is interesting that the students with lower counts are
+mostly above this line and those with higher counts are all below it.
+There is a positive correlation between the pairs of counts,
+consistent with systematic observer effects (i.e., people who count a
+low value in one grain tend to count similarly low in the other, and
+vice versa). Comparing other pairs of grains shows similar results,
+with correlations varying between 0.4 and 0.9.\medskip
+
+\noindent\begin{minipage}{.55\linewidth}
+  \includegraphics[width=\linewidth]{stripchart.pdf}\medskip
+\end{minipage}
+\begin{minipage}{.45\linewidth}
+\captionof{figure}{ Dot plot of the data shown in
+  Table~\ref{tab:crowdtable}, with the students' medians and PV's
+  counts shown as yellow and blue triangles, respectively.
+  \medskip}
+  \label{fig:stripchart}
+\end{minipage}
 
@@
 (Section~\ref{sec:outlook}).\medskip
-
-\begin{small}
-  \input{../output/shortcrowdtable.txt}\\
-\end{small}
-\captionof{table}{Counts of fission tracks in the same ROIs of 25
-  grains made by PV and 68 students. Only 20 grains and 20 students
-  are shown due to space constraints; the full table is provided in
-  the supplementary information.  Each student counted a subset of the
-  grains. The grains are listed in increasing order of PV's
-  counts. The last 3 columns give, for each student, the number of
-  grains counted $n_g$ and the weighted mean $r_p =
-  \sum\,w_i\,r_i/\sum\,w_i$ and standard deviation $s_p =
-  \sqrt{w_i\,(r_i-r_p)^2/(n_g − 1)}$ where $r_i$ is the ratio of the
-  student's count to PV's count for grain $i$ and $w_i$ is PV's count
-  for grain $i$. ($r_p$ is the same as the ratio of the student's
-  total count over the $n_g$ grains to PV's total count over the same
-  grains.)  The students are listed in decreasing order of $r_p$ . The
-  last 3 rows ($n$, $\bar{y}$ and $s_y$ ) give the sample size (number
-  of students), mean and standard deviation of the students' counts
-  for each grain.}
-\label{tab:crowdtable}
-%\end{minipage}
 

### 2025-09-28 geochron@home/R/helper.R

@@ -240,7 +240,10 @@ compare_grains <- function(results,grain1,grain2,plot=TRUE,...){
         bg <- rep(NA,ns)
         iadmin <- which(users==1)
         bg[iadmin] <- 'blue'
-        plot(out$x,
+        duplo <- duplicated(out$x)
+        X <- out$x
+        X[duplo,] <- jitter(X[duplo,])
+        plot(X,
              xlab=paste0('grain ',grain1),
              ylab=paste0('grain ',grain2),
              pch=21,bg=bg,...)
@@ -341,24 +344,22 @@ counts2latex <- function(lst,destination,short=FALSE){
     sy <- sprintf('%.1f',lst$sy)
     short_ybar <- gsub('0\\.', '.', ybar)
     short_sy <- gsub('0\\.', '.', sy)
-    medians <- apply(lst$tab,2,median,na.rm=TRUE)
+    medians <- apply(lst$tab[-1,],2,median,na.rm=TRUE)
+    r_m <- sprintf('%.2f',medians/lst$tab[1,])
+    short_rm <- gsub('0\\.', '.',r_m)
     out <- rbind(c('user','~',colnames(lst$tab),'~','$n_g$','$r_p$','$s_p$'),
                  c('~','~',rep(NA,ncol(lst$tab)),'~','~','~','~'),
                  c('PV',NA,lst$tab[1,],NA,ng[1],short_rp[1],short_sp[1]),
                  NA,
                  cbind(rownames(lst$tab)[-1],NA,lst$tab[-1,],NA,
                        ng[-1],short_rp[-1],short_sp[-1]),
-                 NA,
-                 c('$n$','~',n,NA,'~','~','~'),
-                 c('$\\bar{y}$','~',short_ybar,NA,'~','~','~'),
-                 c('$s_y$','~',short_sy,NA,'~','~','~'))
+                 c('$r_m$','~',short_rm,'~','~','~','~'))
     if (short){
-        out <- out[-(15:(nrow(out)-15)),-(13:(ncol(out)-15))]
-        out[,13] <- '$\\ldots$'
-        out[15,] <- '$\\vdots$'
-        out[c(2,4,15,15+11),13] <- '~'
-        out[15,c(2,13,13+11)] <- '~'
-        out[15,13] <- '$\\ddots$'
+        out <- out[-(15:(nrow(out)-11)),]
+        out[15,] <- '~'
+        out[15,1] <- "$\\vdots$"
[…]

### 2025-09-28 geochron@home/R/main.R

@@ -31,7 +31,7 @@ PAsPA <- matrix(NA,nrow=ng,ncol=2,
                 dimnames=list(grains,c('PA','sPA')))
 NA120 <- matrix(NA,nrow=ng,ncol=7,
                 dimnames=list(1:ng,c('grain','N1','A1','N2','A2','N0','A0')))
-rhoP <- rhoA <- rep(NA,ng)
+rhoP <- rhoA <- w <- r <- rep(NA,ng)
 for (i in seq_along(grains)){
     grain <- grains[i]
     ROIP <- PieterResults[[grain]]$ROI
@@ -45,6 +45,8 @@ for (i in seq_along(grains)){
     xyA <- AndyResults[[grain]]$counts
     xyPinA0 <- xyP[apply(xyP,1,point_in_polygon,ROI0),]
     xyAinA0 <- xyA[apply(xyA,1,point_in_polygon,ROI0),]
+    w[i] <- nrow(xyPinA0)
+    r[i] <- nrow(xyAinA0)/nrow(xyPinA0)
     N1 <- nrow(xyP)
     N2 <- nrow(xyA)
     total_counts <- total_counts + c(N1,N2)
@@ -57,6 +59,8 @@ for (i in seq_along(grains)){
     PAsPA[i,'sPA'] <- PAsPA[i,'PA'] * sqrt(1/N1+1/N2-2*N00/(N1*N2))
     NA120[i,] <- c(grain,N1,A1,N2,A2,N00,A0)
 }
+rp <- sum(w*r)/sum(w)
+sp <- sqrt(sum(w*(r-rp)^2/(ng-1)))
 
 write.table(NA120,file='../output/PA.csv',sep=',')
 
@@ -152,21 +156,6 @@ lines(x=lim,y=lim,col='grey50',lwd=1.5)
 par(op)
 dev.off()
 
-#### 6. Boxplot of crowdsourcing results ####
-pdf(file='../output/boxplots.pdf',width=6,height=5)
-op <- par(mar=c(3,3,1,8),mgp=c(2,1,0))
-# box plot
-grouped_list <- split(trustworthy_results$count, trustworthy_results$index)
-boxplot(grouped_list,horizontal=TRUE,
-        xlab='Ns',ylab='grain number',
-        xaxt='n',las=2,col=NA)
-add_admin_count_to_boxplot(trustworthy_results,grouped_list)
-axis(1)
-# table
-add_table(grouped_list)
-par(op)
-dev.off()
-
[…]

### 2025-09-28 geochron@home/manuscript/paper.tex
Updated Copernicus style sheets.

@@
 that was carried out at UCL as part of an undergraduate course in
-isotope geology. 70 students were asked to create an account on the
+isotope geology. 68 students were asked to create an account on the
 geochron@home server by selecting a unique username and
@@
 
-The raw counts for 68 students (two of the students were unable to
-complete the exercise) are given in Table~\ref{tab:crowdtable} and
-Figure~\ref{fig:stripchart}, for the same selection of 25 grains that
-were analysed by AC and PV in Section~\ref{sec:private}, along with
-corresponding counts made by PV. Everyone counted tracks in the same
-ROIs. Variation between the mean values is to be expected because of
-the differing numbers of tracks due to varying areas and U contents of
-the grains. But the variation between counts within grains (shown by
-the standard deviations in Figure~\ref{fig:stripchart}) is due
-entirely to differences between the students' recognising and counting
-exactly the same tracks. These standard deviations increase with the
-mean number of tracks and are considerable in size, being on average
-about 40\% of the mean. With expert trained counters one would expect
-much smaller differences between counts.  Furthermore, the vast
-majority of students counted fewer tracks than PV did, often many
-fewer, and on two occasions someone counted no tracks at all. PV's
-count is always above the students' median and nearly always above the
-upper quartile (Figure~\ref{fig:stripchart}).
-Table~\ref{tab:crowdtable} (looking along rows) shows just 7 students
-with average counts within 10\% of PV's, and 18 with average counts
-less than half of his.\medskip
+The raw counts for the 68 students are given in
+Table~\ref{tab:crowdtable} and Figure~\ref{fig:stripchart}, for the
+same selection of 25 grains that were analysed by AC and PV in
+Section~\ref{sec:private}, along with corresponding counts made by
+PV. Everyone counted tracks in the same ROIs. Variation between the
+mean values is to be expected because of the differing numbers of
+tracks due to varying areas and U contents of the grains. But the
+variation between counts within grains (shown by the standard
+deviations in Figure~\ref{fig:stripchart}) is due entirely to
+differences between the students' recognising and counting exactly the
+same tracks. These standard deviations increase with the mean number
+of tracks and are considerable in size, being on average about 40\% of
+the mean. With expert trained counters one would expect much smaller
+differences between counts.  Furthermore, the vast majority of
+students counted fewer tracks than PV did, often many fewer, and on
+two occasions someone counted no tracks at all. PV's count is always
+above the students' median and nearly always above the upper quartile
+(Figure~\ref{fig:stripchart}).\medskip
 
 The rightmost three columns of Table~\ref{tab:crowdtable} contain
-summary statistics ($n_g$ , $r_p$ and $s_p$) giving, for each student,
+summary statistics ($n_g$, $r_p$ and $s_p$) giving, for each student,
 the number of grains counted and the weighted mean and standard
@@
 $r_p = 1$ and $s_p = 0$. For comparison, an $r_p = 1.04$ and $s_p =
-0.69$ for PV and AC using just the counts in the common areas $A_0$
-from Section~\ref{sec:private}. With the caveat that the ROIs are
-different and PV's counts are different, this calculation shows that
-there is a group of 6 (possibly 9) students whose agreement with PV is
-comparable with AC's. This group does not include student 229 whose
-$r_p = 0.95$ but with a large $s_p = 1.62$.\medskip
+0.69$ is obtained for PV and AC using just the counts in the common
+areas $A_0$ from Section~\ref{sec:private}. With the caveat that the
+ROIs are different and PV's counts are different, this calculation
+shows that there is a group of 6 (possibly 9) students whose agreement
+with PV is comparable with AC's. This group does not include student
+229 whose $r_p = 0.95$ but with a large $s_p = 1.62$.\medskip
 
-However, in spite of the under-counting by the students, the ratio of
-their pooled track counts ($r_p$) for two different grains is often
-close to PV's ratio, especially when comparing grains with similar
-track densities. This is illustrated in Figure~\ref{fig:23vs25}a for
+In spite of the under-counting by the students, the ratio of their
+pooled track counts ($r_p$) for two different grains is often close to
+PV's ratio, especially when comparing grains with similar track
+densities. This is illustrated in Figure~\ref{fig:23vs25}a for
 grains~23 and 25.  For any point, the ratio of the two counts is given

### 2025-09-28 geochron@home/R/helper.R

@@ -355,7 +355,7 @@ counts2latex <- function(lst,destination,short=FALSE){
                        ng[-1],short_rp[-1],short_sp[-1]),
                  c('$r_m$','~',short_rm,'~','~','~','~'))
     if (short){
-        out <- out[-(15:(nrow(out)-11)),]
+        out <- out[-(15:(nrow(out)-12)),]
         out[15,] <- '~'
         out[15,1] <- "$\\vdots$"
         out <- out[-c(2,4),]

### 2025-10-02 geochron@home/manuscript/paper.tex
Flipped numerator and denominator in grain 1 vs 2 comparison

@@
 individual slopes vary considerably, the line with slope equal to the
-ratio of the total counts for all students (which is 0.96) passes
+ratio of the total counts for all students (which is 0.97) passes
 close to PV's pair of counts (52,50) shown by the blue dot (with ratio
-1.03). It is interesting that the students with lower counts are
+1.04). It is interesting that the students with lower counts are
 mostly above this line and those with higher counts are all below it.
@@
 \begin{minipage}{.45\linewidth}
-\captionof{figure}{ Dot plot of the data shown in
+\captionof{figure}{Dot plot of the data shown in
   Table~\ref{tab:crowdtable}, with the students' medians and PV's
@@
     those for grain 25 for 36 students who counted both of those
-    grains. The grey line through the origin has slope equal to the
-    ratio of the total counts for each grain. The blue dot shows PV's
-    counts for those two grains. b) Optical images in transmitted
-    light of Mount Dromedary apatites 23 and 25 in the crowd-sourcing
-    experiment, along with two-dimensional histogram of the track
-    locations for the two grains, as identified by the citizen
-    scientists.
+    grains. The grey line through the origin has slope 0.97, equal to
+    the ratio of the students' total counts for each grain.  The blue
+    dot shows PV's counts for those two grains. b) Optical images in
+    transmitted light of Mount Dromedary apatites 23 and 25 in the
+    crowd-sourcing experiment, along with two-dimensional histogram of
+    the track locations for the two grains, as identified by the
+    citizen scientists.
     \medskip}

### 2025-10-02 geochron@home/R/helper.R

@@ -222,19 +222,19 @@ add_admin_count_to_boxplot <- function(trustworthy_results,grouped_list){
     points(admin_counts,1:length(ACids),pch=21,bg='blue')
 }
 
-compare_grains <- function(results,grain1,grain2,plot=TRUE,...){
-    results1 <- results[results$index == grain1,]
-    results2 <- results[results$index == grain2,]
-    users <- intersect(results1$user_id,
-                       results2$user_id)
-    i1 <- match(users,results1$user_id)
-    i2 <- match(users,results2$user_id)
-    out <- list(name=paste0(grain1,' vs ',grain2))
+compare_grains <- function(results,grain_x,grain_y,plot=TRUE,...){
+    results_x <- results[results$index == grain_x,]
+    results_y <- results[results$index == grain_y,]
+    users <- intersect(results_x$user_id,
+                       results_y$user_id)
+    i_x <- match(users,results_x$user_id)
+    i_y <- match(users,results_y$user_id)
+    out <- list(name=paste0(grain_x,' vs ',grain_y))
     class(out) <- 'counts'
-    out$x <- cbind(results1$count[i1],
-                   results2$count[i2])
-    colnames(out$x) <- c(paste0('Ns(',grain1,')'),
-                         paste0('Ns(',grain2,')'))
+    out$x <- cbind(results_x$count[i_x],
+                   results_y$count[i_y])
+    colnames(out$x) <- c(paste0('Ns(',grain_x,')'),
+                         paste0('Ns(',grain_y,')'))
     ns <- nrow(out$x)
     if (plot){
         bg <- rep(NA,ns)
@@ -244,17 +244,17 @@ compare_grains <- function(results,grain1,grain2,plot=TRUE,...){
         X <- out$x
         X[duplo,] <- jitter(X[duplo,])
         plot(X,
-             xlab=paste0('grain ',grain1),
-             ylab=paste0('grain ',grain2),
+             xlab=paste0('grain ',grain_x),
+             ylab=paste0('grain ',grain_y),
              pch=21,bg=bg,...)
         rho <- format(round(cor(out$x)[1,2],2),nsmall=2)
         mtext(line=0,text=paste0('correlation=',rho),cex=1.0)
-        PVratio <- out$x[iadmin,1]/out$x[iadmin,2]
+        PVratio <- out$x[iadmin,2]/out$x[iadmin,1]
         mtext(line=-2,
               text=paste0("PV's ratio=",
[…]

### 2025-10-02 geochron@home/R/main.R

@@ -151,7 +151,7 @@ trustworthy_results <- clean_results(results)
 pdf(file='../output/crowd-correlation.pdf',width=5,height=5)
 op <- par(mar=c(3,3,1,1),mgp=c(2,1,0),xpd=NA,bty='n')
 lim <- c(0,70)
-compare_grains(trustworthy_results,grain2$index,grain1$index,xlim=lim,ylim=lim)
+compare_grains(trustworthy_results,grain_x=grain2$index,grain_y=grain1$index,xlim=lim,ylim=lim)
 lines(x=lim,y=lim,col='grey50',lwd=1.5)
 par(op)
 dev.off()

### 2025-10-03 geochron@home/manuscript/paper.tex
G@H manuscript ready for final round of internal review.

@@
   grains made by PV and 68 students. Only 20 students are shown due to
-  space constraints; the full table is provided in the supplementary
-  information.  Each student counted a subset of the grains. The
-  grains are listed in increasing order of PV's counts. The last 3
-  columns give, for each student, the number of grains counted $n_g$
-  and the weighted mean $r_p = \sum\,w_i\,r_i/\sum\,w_i$ and standard
-  deviation $s_p = \sqrt{\sum w_i\,(r_i-r_p)^2/(n_g − 1)}$
-  where $r_i$ is the ratio of the student's count to PV's count for
-  grain $i$ and $w_i$ is PV's count for grain $i$. ($r_p$ is the same
-  as the ratio of the student's total count over the $n_g$ grains to
-  PV's total count over the same grains.)  The students are listed in
-  decreasing order of $r_p$. $r_m$ is the ratio of the students'
-  median count per grain to PV's count.}
+  space constraints; the full table is provided in the GaHa.  Each
+  student counted a subset of the grains. The grains are listed in
+  increasing order of PV's counts. The last 3 columns give, for each
+  student, the number of grains counted $n_g$ and the weighted mean
+  $r_p = \sum\,w_i\,r_i/\sum\,w_i$ and standard deviation $s_p =
+  \sqrt{\sum w_i\,(r_i-r_p)^2/(n_g − 1)}$ where $r_i$ is the ratio of
+  the student's count to PV's count for grain $i$ and $w_i$ is PV's
+  count for grain $i$. ($r_p$ is the same as the ratio of the
+  student's total count over the $n_g$ grains to PV's total count over
+  the same grains.)  The students are listed in decreasing order of
+  $r_p$. $r_m$ is the ratio of the students' median count per grain to
+  PV's count.}
 \label{tab:crowdtable}
@@
 \begin{minipage}{.45\linewidth}
-\captionof{figure}{Dot plot of the data shown in
+\captionof{figure}{Strip chart of the data shown in
   Table~\ref{tab:crowdtable}, with the students' medians and PV's

### 2025-10-03 geochron@home/R/main.R

@@ -74,7 +74,7 @@ for (grain in grains){
 par(op)
 dev.off()
 
-#### 3. Compare PV and AC's results for grain 3 (a.k.a. 4648) ####
+#### 3. Fig 2: Compare PV and AC's results for grain 3 (a.k.a. 4648) ####
 pdf(file='../output/AvP.pdf',width=15,height=5)
 op <- par(mar=c(4,4,.5,.5),mfrow=c(1,3),mgp=c(2.5,1,0),cex=1.0)
 plotROIs(PieterResults[['4648']],AndyResults[['4648']])
@@ -112,9 +112,9 @@ for (grain in grains){
 }
 dev.off()
 
-#### 5. Show crowdsourcing results for two selected grains ####
-grain1 <- list(number='4685',index=23) # list(number='4682',index=39)
-grain2 <- list(number='4674',index=25) # list(number='4680',index=43)
+#### 5. Fig. 4b: Show crowdsourcing results for two selected grains ####
+grain1 <- list(number='4685',index=23)
+grain2 <- list(number='4674',index=25)
 pdf(file="../output/23vs25.pdf",onefile=FALSE)
 op <- par(mar=c(4,4,0,1))
 image1 <- plotimage(idir='../screenshots',
@@ -145,27 +145,22 @@ grid.arrange(
 par(op)
 dev.off()
 
-trustworthy_results <- clean_results(results)
-
-#### 6. Scatter plot of crowdsourcing results ####
+#### 6. Fig. 4a: Scatter plot of crowdsourcing results ####
 pdf(file='../output/crowd-correlation.pdf',width=5,height=5)
 op <- par(mar=c(3,3,1,1),mgp=c(2,1,0),xpd=NA,bty='n')
 lim <- c(0,70)
+trustworthy_results <- clean_results(results)
 compare_grains(trustworthy_results,grain_x=grain2$index,grain_y=grain1$index,xlim=lim,ylim=lim)
 lines(x=lim,y=lim,col='grey50',lwd=1.5)
 par(op)
 dev.off()
 
-# generalised linear fit to crowd-sourced data
-fit <- glm(count ~ user_id + index, data=trustworthy_results, family="poisson")
-mswd <- summary(fit)$deviance/summary(fit)$df.residual
-
-# summary tables
+#### 7. Tab. 1: summary tables ####
 lst <- crowdtable(trustworthy_results)
[…]

### 2025-10-03 geochron@home/manuscript/paper.tex
Mentioned student vandalism

@@
 with PV is comparable with AC's. This group does not include student
-229 whose $r_p = 0.95$ but with a large $s_p = 1.62$.\medskip
+229 whose $r_p = 0.95$ but with a large $s_p = 1.62$. Students near
+the bottom of Table~\ref{tab:crowdtable} have undercounted the samples
+by such a large degree that their work can be qualified as
+vandalism. Section~\ref{sec:outlook} will suggest strategies to detect
+and counteract this type of behaviour.\medskip
 

### 2025-10-04 geochron@home/manuscript/paper.tex
Incorporated AC's suggestions

@@
 The combined data plots of Figures~\ref{fig:AvP}b and c contain two
-sets of counts, with $\sum N_{1} = 686$ and $\sum N_{2} = 679$, and a
-weighted mean track density ratio $\hat{\rho}_{1}/\hat{\rho}_{1} =
-0.94$ with relative standard error 0.013. PV counted 600 tracks in
-those common areas $A_0$, of which 549 were also counted by
-AC. Conversely, AC counted 622 tracks of which 549 were also counted
-by PV. The ratio of the two analysts' track density estimates based
-just on the common area is therefore 600/622 = 0.965 with relative
-standard error 0.018, calculated from Equation~\ref{eq:NO}, with $N_1
-= 686$, $N_2 = 679$ and $N_0 = 549$.  This is close to the weighted
-mean ratio of 0.94 (Figure~\ref{fig:AvP}c), and is slightly nearer to
-1. It indicates that PV under-counts the Mount Dromedary apatite by
-3.5\% relative to AC.  The existence of `observer bias' is well
-documented \citep{tamer2025}.  It is one of the reasons why fission
-track analysis is often done relative to age standards: observer bias
-does not have to be a problem provided that it is consistent between
+sets of counts for 25 grains, with $\sum N_{1} = 686$ (PV) and $\sum
+N_{2} = 679$ (AC), and a weighted mean track density ratio
+$\hat{\rho}_{1}/\hat{\rho}_{1} = 0.94$ with relative standard error
+0.013. PV counted 600 tracks in those common areas $A_0$, of which 549
+were also counted by AC. Conversely, AC counted 622 tracks of which
+549 were also counted by PV. The ratio of the two analysts' track
+density estimates based just on the common area is therefore 600/622 =
+0.965 with relative standard error 0.018, calculated from
+Equation~\ref{eq:NO}, with $N_1 = 686$, $N_2 = 679$ and $N_0 = 549$.
+This is close to the weighted mean ratio of 0.94
+(Figure~\ref{fig:AvP}c), and is slightly nearer to 1. It indicates
+that PV under-counts the Mount Dromedary apatite by 3.5\% relative to
+AC.  The existence of `observer bias' is well documented
+\citep{tamer2025}.  It is one of the reasons why fission track
+analysis is often done relative to age standards: observer bias does
+not have to be a problem provided that it is consistent between
 grains, and between samples.\medskip
@@
 
-The annotations in the current tutorial pages were made by an
-experienced fission track analyst (Andrew Carter). This basic tutorial
-provides a quick and easy mechanism to train novice users in the art
-of fission track analysis. The tutorial pages are in their infancy
-and offer only a limited degree of interactivity. Users can click on
-features to read the annotations. The tutorial will grow with input
-from experienced analysts and future plans include the addition of
-fully interactive `quizzes' (Section~\ref{sec:conclusions}). The
-limitations of the current tutorials are apparent in the results of
-the crowd-sourcing experiment described in the next section of this
-paper.
+Some exemplar annotations in the current tutorial pages were made by
+an experienced fission track analyst (Andrew Carter). This basic
+tutorial provides a quick and easy mechanism to train novice users in
+the art of fission track analysis. The tutorial pages are in their
+infancy and offer only a limited degree of interactivity. Users can
+click on features to read the annotations. A more detailed,
+comprehensive tutorial will grow with the wider input from experienced
+analysts and future plans include the addition of fully interactive
+`quizzes' (Section~\ref{sec:conclusions}). The limitations of the
+current tutorials are apparent in the results of the crowd-sourcing
+experiment described in the next section of this paper.
 
@@
 geochronological developments.  Because fission track data are
-imprecise, the fission track community has sollicited the help of
+imprecise, the fission track community has solicited the help of
 statisticians and mathematicians to develop its analytical protocols.

### 2025-10-06 geochron@home/manuscript/paper.tex
Fixed spelling

@@
 of charge. The archive accepts image stacks acquired on any type of
-digital microscope, and accomodates fission track data (counts and
+digital microscope, and accommodates fission track data (counts and
 length measurements) from external fission track analysis suites such
